---
title: "phosphoDataset msqrobPTM without GP"
output:
  html_document:
    code_download: yes
    theme: cosmo
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
    highlight: tango
    number_sections: yes
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: '4'
---

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "C:/Users/Nina/OneDrive - UGent/Documenten/Doctoraat/msqrobPTM paper/biological phospho dataset/")
```

```{r message=FALSE}
library(QFeatures)
library(msqrob2)
library(dplyr)
library(tidyr)
library(tibble)
library(gt)
library(plotly)
library(stageR)
library(poolr)
library(RColorBrewer)
library(seqinr)
library(stringr)
library(ExploreModelMatrix)
library(data.table)
```



# Import data


```{r}
df <- read.csv("evidence.txt",header=TRUE, sep = ",")
#I can ignore the sigma proteins (those were just used for QC purposes)
df <- df %>% filter(!grepl("Sigma", df$Proteins))
#filter out phospho sites with less than 75% probability
#(other filtering options are possible)
filter_rows <- sapply(df$Phospho..STY..Probabilities, function(x){
  if (!x==""){
    prob <- str_extract_all(x,  "(?<=\\().+?(?=\\))")[[1]]
    if (!any(as.double(prob) > 0.75)){return(which(df$Phospho..STY..Probabilities == x))}
  }
}, USE.NAMES = F)
filter_rows <- unique(unlist(filter_rows))
nrow(df)
df <- df[-filter_rows,]
nrow(df)
#Now get format into wide format
df_wide <- pivot_wider(df, id_cols = c("Sequence", "Modifications", "Modified.sequence", "Proteins", "Leading.proteins",
                                        "Reverse", "Potential.contaminant", "Protein.group.IDs", "Leading.razor.protein"),
                                                                          #Is dat hier de beste manier om met 
                                                                          #gedupliceerde features om te gaan?
                       names_from = "Raw.file", names_prefix = "Intensity_", values_from = "Intensity", values_fn = max)
```


```{r}
ecols <- grep("Intensity",colnames(df_wide))
#order dataframe by protein, for the normalisation step
df_wide = df_wide[order(df_wide$Leading.razor.protein),]
pe <- readQFeatures(df_wide,ecol= ecols,name="peptidoformRaw")
```

```{r}
rownames(pe[["peptidoformRaw"]]) <- rowData(pe[["peptidoformRaw"]])$Modified.sequence
```

## Experimental Layout

```{r}
metadata <- read.csv("Experimental Design.csv")
```

```{r}
colData(pe)$file <- sapply(str_split(rownames(colData(pe)), "_"), function(x) x[[2]])
metadata$File <- sub(x=metadata$File, pattern=".raw",replacement = "")
colData(pe)$condition <- sapply(colData(pe)$file, function(x){
  metadata[metadata$File==x,]$Condition
}, USE.NAMES = F)
colData(pe)$subset <- sapply(colData(pe)$file, function(x){
  metadata[metadata$File==x,]$Subset
}, USE.NAMES = F)
colData(pe)$condition <- case_when(grepl("A", colData(pe)$condition) ~ "A",
                                        TRUE ~ "B")
colData(pe)
```


```{r}
MSnbase::plotNA(assay(pe[["peptidoformRaw"]])) +
  xlab("Peptide index (ordered by data completeness)")
```


# Preprocessing


```{r}
rowData(pe[["peptidoformRaw"]])$nNonZero <- rowSums(assay(pe[["peptidoformRaw"]]) > 0, na.rm = T)
pe <- zeroIsNA(pe, i = "peptidoformRaw")
pe <- filterFeatures(pe, ~Reverse != "+")
pe <- filterFeatures(pe, ~Potential.contaminant != "+")
pe <- logTransform(pe, base = 2, i = "peptidoformRaw", name = "peptidoformLog")
pe <- pe[rowData(pe[["peptidoformRaw"]])$nNonZero>2,,]
pe <- QFeatures::normalize(pe, method = "center.median", i = "peptidoformLog", name = "peptidoform")
colData(pe[["peptidoform"]]) <- colData(pe)
```


## Normalisation via robust summarisation


```{r,warning=FALSE}
pe <- aggregateFeatures(pe,
 i = "peptidoform",
 fcol = "Leading.razor.protein",
 na.rm = TRUE,
 name = "proteinRobust",
 fun = MsCoreUtils::robustSummary)
```

```{r}
#Only take pepforms that have a parent protein
#(when there is no global profiling dataset, this will be all peptidoforms)
pepWithProtein <- rowData(pe[["peptidoform"]])$Proteins %in% rownames(pe[["proteinRobust"]])
pePepWithProtein  <- pe[["peptidoform"]][pepWithProtein,]
pe <- addAssay(pe,pePepWithProtein,"pepformRel")
#normalisation for protein abundance step
assay(pe[["pepformRel"]]) <- assay(pe[["pepformRel"]]) - assay(pe[["proteinRobust"]])[rowData(pe[["pepformRel"]])$Proteins,colnames(assay(pe[["pepformRel"]]))]
boxplot(assay(pe[["pepformRel"]]))
```


# Differential peptidoform usage (DPFU)

## Hypothesistest for each contrast

```{r}
colData(pe)$condition <- relevel(as.factor(colData(pe)$condition), ref = "B")
colData(pe)$subset <- as.factor(colData(pe)$subset)
```

```{r}
pe <- msqrob2::msqrob(object = pe, i = "pepformRel", formula = ~condition*subset, robust=FALSE)
rowData(pe[["pepformRel"]])$msqrobModels[[2]] %>%
                  getCoef
```


```{r}
fraction_of_ys <- (colData(pe) %>% as_tibble() %>% filter(subset == "y") %>% nrow()) / nrow(colData(pe))
contrasts <- c("conditionA", "conditionA + conditionA:subsety", "conditionA:subsety", "conditionA + 0.5 * conditionA:subsety", "conditionA + 0.6666667 * conditionA:subsety")
L <- makeContrast(c("conditionA = 0",
                    "conditionA + conditionA:subsety = 0",
                    "conditionA:subsety = 0",
                    "conditionA + 0.5 * conditionA:subsety = 0",
                    "conditionA + 0.6666667 * conditionA:subsety = 0"),
                  parameterNames = rowData(pe[["pepformRel"]])$msqrobModels[[2]] %>%
                  getCoef %>%
                  names)
pe <- hypothesisTest(object = pe, i = "pepformRel", contrast = L, overwrite = T)
```


### Volcanoplot 

```{r}
volcano <- list()
for (contrast in contrasts){
volcano[[contrast]] <- rowData(pe[["pepformRel"]])[[contrast]]%>%
            ggplot(aes(x = logFC, y = -log10(pval), 
                   color = adjPval < 0.05,
                   annotation=rowData(pe[["pepformRel"]])[,3])) +
  geom_point(cex = 2.5) +
  scale_color_manual(values = alpha(c("black", "red"), 0.5)) + 
  theme_minimal() +
  ylab("-log10(pvalue)") +
  ggtitle(contrast)
}
volcano
```

```{r eval=FALSE, fig.height=4, fig.width=12, include=FALSE}
for(contrast in contrasts){
sigNames <- rowData(pe[["pepformRel"]])[[contrast]] %>%
 rownames_to_column("peptidoform") %>%
 filter(adjPval<0.05) %>%
 pull(peptidoform)
heatmap(assay(pe[["pepformRel"]])[sigNames, ], main = contrast)
}
```

### Significant peptidoforms for each contrast


```{r}
tables <- list()
for (contrast in contrasts){
sigTable <- rowData(pe[["pepformRel"]])[[contrast]]
if(nrow(sigTable %>%
  na.exclude %>%
  filter(adjPval<0.05))>0){
sigTable <- sigTable %>%
  na.exclude %>%
  filter(adjPval<0.05) %>%
  arrange(pval) %>%
  mutate(
    se = format(se, digits = 2), 
    df = format(df, digits =2),
    t = format(t, digits = 2),
    adjPval = format(adjPval, digits = 3),
    rank = 1:length(logFC) 
  ) 
sigTable_print <- sigTable %>% mutate(peptidoform = rownames(sigTable)) %>% gt() %>% tab_header(title = md(contrast))
tables[[contrast]] <- sigTable
}
}
knitr::kable(tables)
```

```{r eval=FALSE, include=FALSE}
tables_toCsv <- do.call("rbind", tables)
tables_toCsv <- tables_toCsv %>% rownames_to_column("contrast")
pepforms <- sapply(unique(tables_toCsv$contrast), function(x){rownames(tables[[x]])}, USE.NAMES = F)
tables_toCsv$Peptidoform <- pepforms
write.csv(tables_toCsv, file = "C:/Users/Nina/OneDrive - UGent/Documenten/Doctoraat/msqrobPTM/PhosphoDataset /significance_tables_peptidoform.csv")
```


# DPTM

## ptm summarisatie

### Determine location of ptm

```{r}
fasta <- "human.fasta"
parsed_fasta <- read.fasta(file = fasta, seqtype = "AA", as.string = T)
```

  
```{r}
#modified sequence column contains _ that does not matter, but hinders location determining
rowData(pe[["pepformRel"]])$modified_sequence <- gsub("_", "", rowData(pe[["pepformRel"]])$Modified.sequence)
```

  

```{r}
get_ptm_location <- function(feature, data, fasta, mod_column = "Modifications", 
                             peptide_seq_column = "Sequence", mod_seq_column = "modified_sequence", 
                             protein_column = "Leading.razor.protein", split = ",", collapse = ", "){
  prot <- data[feature,][[protein_column]]
  pep_seq <- data[feature,][[peptide_seq_column]]
  mod_seq <- data[feature,][[mod_seq_column]]
  prot_seq <- fasta[[prot]][1]

  #find location of peptide in protein
  #add -1 here, so that the addition of the location later on is correct
  pep_location <- unlist(lapply(gregexpr(pattern = pep_seq, prot_seq), min)) - 1
  final_mod <- c()
  j <- mod_seq
  #go over the modifications inside the modified sequence
  for(mod in regmatches(mod_seq, gregexpr("\\(.*?\\)\\)", mod_seq, perl=T))[[1]]){
    #find location of modification in peptide
    mod_location <- unlist(lapply(gregexpr(pattern = mod, j, fixed = T), min))
    #get location in protein (-1 because else it gives you the location after because of the presence of the modification in the string)
    location <- mod_location + pep_location -1
    #add location to modification
    mod_ <- paste(mod, location)
    #save modification
    final_mod <- c(final_mod, mod_)
    #now remove current modification from the sequence, so that we can continue to the next mod
    str_sub(j, mod_location, nchar(mod)+mod_location-1) <- ""
  }
  return(paste(final_mod, collapse = collapse))
}
```

```{r}
rowData(pe[["pepformRel"]])$mod <- sapply(rownames(rowData(pe[["pepformRel"]])), function(x){
    get_ptm_location(x, rowData(pe[["pepformRel"]]), parsed_fasta)
}, USE.NAMES = F)
```


```{r}
#Add ptm variable = protein + modification
rowData(pe[["pepformRel"]])$ptm <- ifelse(rowData(pe[["pepformRel"]])$mod != "",
                                               paste(rowData(pe[["pepformRel"]])$Leading.razor.protein, rowData(pe[["pepformRel"]])$mod, sep="_"), 
                                               "")
```

### Get ptm level intensity matrix

Get all unique ptms present in the dataset (all protein - modification - location combinations)

```{r}
prots <- unique(rowData(pe[["pepformRel"]])$Leading.razor.protein)
#Do for each protein
ptms <- sapply(prots, function(i) {
  pe_sub <- pe[["pepformRel"]][grepl(i, rowData(pe[["pepformRel"]])$Leading.razor.protein, fixed = T),]
  #pe_sub <- filterFeatures(pe,~grepl(Leading.razor.protein,pattern=i,fixed = T))
  #Get all unique modification present on that protein
  mods <- unique(unlist(strsplit(rowData(pe_sub)$mod, split = ", ", fixed = TRUE)))
  #Add protein info to mods
  ptm <- paste(rep(i, length(mods)), mods)
  #return all the protein-mods combinations
  ptm
})
ptms <- as.vector(unlist(ptms))
```



```{r}
#For each ptm do
ptm_x_assay <- sapply(seq(1:length(ptms)), function(i){ 
  x <- ptms[i]
  #Get current protein and mod from ptm
  prot <- str_split(x, " ", 2)[[1]][1]
  current_ptm <- str_split(x, " ", 2)[[1]][2]
  #filter on that protein and on that mod to obtain all peptidoforms that correspond to the ptm
  pe_sub <- pe[["pepformRel"]][grepl(prot, rowData(pe[["pepformRel"]])$Leading.razor.protein, fixed = T),]
  #pe_sub <- filterFeatures(pe,~grepl(Leading.razor.protein,pattern=prot, fixed = T))
  ptm_sub <- pe_sub[grepl(current_ptm, rowData(pe_sub)$mod, fixed = T),]
  #ptm_sub <- filterFeatures(pe_sub,~grepl(mod,pattern=current_ptm, fixed=T))[["peptidoformNorm"]]
  #Get intensity values of those peptidoforms
  y <- assay(ptm_sub)
  #And summarise them to 1 row of intensity values: 1 value per sample for that ptm
  ptm_y <- try(MsCoreUtils::robustSummary(y), silent = T)
  if (is(ptm_y, "try-error")){
    ptm_y <- rep(NA, ncol(y))}
  ptm_y
})
#Then we get the intensity assay on ptm level
ptm_x_assay <- t(ptm_x_assay)
rownames(ptm_x_assay) <- ptms
```

### Add to QFeatures object

Filter out ptms with all zero intensities

```{r}
print(paste(nrow(ptm_x_assay), "ptms before filtering"))
filtering <- rowSums(ptm_x_assay != 0, na.rm=TRUE) > 0 
ptm_x_assay <- ptm_x_assay[filtering,]
print(paste(nrow(ptm_x_assay), "ptms after filtering"))
```

```{r}
all(rownames(colData(pe)) == colnames(ptm_x_assay))
rowdata <- data.frame(ptm = rownames(ptm_x_assay))
rowdata$protein <- sapply(str_split(rowdata$ptm, pattern=" "),function(x) x[1])
ptm_y_assay <- SummarizedExperiment(assays=as.matrix(ptm_x_assay), rowData=rowdata, colData=colData(pe))
```


```{r}
pe <- addAssay(pe, ptm_y_assay, "ptmRel")
```



## Hypothesistest for each contrast


```{r}
pe <- msqrob2::msqrob(object = pe, i = "ptmRel", formula = ~condition*subset, robust=TRUE, overwrite = T)
rowData(pe[["ptmRel"]])$msqrobModels[[2]] %>%
                  getCoef
```



```{r}
contrasts <- c("conditionA", "conditionA + conditionA:subsety", "conditionA:subsety", "conditionA + 0.5 * conditionA:subsety", "conditionA + 0.6666667 * conditionA:subsety")
L <- makeContrast(c("conditionA = 0",
                    "conditionA + conditionA:subsety = 0",
                    "conditionA:subsety = 0",
                    "conditionA + 0.5 * conditionA:subsety = 0",
                    "conditionA + 0.6666667 * conditionA:subsety = 0"),   
                  parameterNames = rowData(pe[["ptmRel"]])$msqrobModels[[2]] %>%
                  getCoef %>%
                  names)
pe <- hypothesisTest(object = pe, i = "ptmRel", contrast = L, overwrite = T)
```



### Volcanoplot 


```{r}
volcanos <- list()
for (contrast in contrasts){
library(plotly)
volcanos[[contrast]] <- 
  rowData(pe[["ptmRel"]])[[contrast]]%>%
  ggplot(aes(x = logFC, y = -log10(pval), 
             color = adjPval < 0.05,
             annotation=rowData(pe[["ptmRel"]])[,3])) +
  geom_point(cex = 2.5) +
  scale_color_manual(values = alpha(c("black", "red"), 0.5)) + theme_minimal() +
  ylab("-log10(pvalue)") +
  ggtitle(contrast)
}
volcanos
```

### Heatmap

We first select the names of the ptms that were declared signficant.

```{r eval=FALSE, include=FALSE}
for(contrast in contrasts){
sigNames <- rowData(pe[["ptmRel"]])[[contrast]] %>%
 rownames_to_column("ptmRel") %>%
 filter(adjPval<0.05) %>%
 pull(ptm)
heatmap(assay(pe[["ptmRel"]])[sigNames, ], main = contrast)
}
```


### Significant ptms for each contrast


```{r}
tables <- list()
for (contrast in contrasts){
sigTable <- rowData(pe[["ptmRel"]])[[contrast]]
if(nrow(sigTable %>%
  na.exclude %>%
  filter(adjPval<0.05)) > 0){
sigTable <- sigTable %>%
  na.exclude %>%
  filter(adjPval<0.05) %>%
  arrange(pval) %>%
  mutate(
    se = format(se, digits = 2), 
    df = format(df, digits =2),
    t = format(t, digits = 2),
    adjPval = format(adjPval, digits = 3),
    rank = 1:length(logFC) 
  ) 
sigTable_print <- sigTable %>% mutate(peptidoform = rownames(sigTable)) %>% gt() %>% tab_header(title = md(contrast))
tables[[contrast]] <- sigTable
}
}
knitr::kable(tables)
```

```{r eval=FALSE, include=FALSE}
tables_toCsv <- do.call("rbind", tables)
tables_toCsv <- tables_toCsv %>% rownames_to_column("rowname")
write.csv(tables_toCsv, file = "significance_tables_PTM.csv")
```


```{r}
a = "sp|Q16610|ECM1_HUMAN (Phospho (STY)) 80"
b = "sp|P13611|CSPG2_HUMAN (Phospho (STY)) 2115"
tables <- list()
for (contrast in contrasts){
sigTable <- rowData(pe[["ptmRel"]])[[contrast]]
tables[[contrast]] <- sigTable %>% rownames_to_column("ptm") %>% filter(ptm==a|ptm==b)
}
knitr::kable(tables)
```

# PLOTS

## Lineplots

### PTM - level

```{r, fig.width=18, fig.height=10}
ptm_list1 <- c()
for (contrast in contrasts){
sigTable <- rowData(pe[["ptmRel"]])[[contrast]] %>% filter(adjPval<0.05)
ptm_list1 <- c(ptm_list1, rownames(sigTable))
}
ptm_list1 <- unique(ptm_list1)

plots <- list()
for (i in ptm_list1){
prot_ = str_split(i, " ")[[1]][1]
site_ = str_split_fixed(i, " ", 2)[,2]

pepform_df <- longFormat(pe[,,"pepformRel"], rowvars = c("Leading.razor.protein", "ptm"), colvars = c("condition", "subset")) %>% as.data.frame()
pepform_df <- pepform_df %>% filter(Leading.razor.protein == prot_)
pepform_df <- pepform_df %>% filter(grepl(site_, ptm, fixed = T))
pepform_df$FeatureType <- "Peptide"
pepform_df <- pepform_df %>% arrange(condition, subset, rowname)

ptm_df <- longFormat(pe[,,"ptmRel"], rowvars = c("protein", "ptm"), colvars = c("condition", "subset")) %>% as.data.frame()
ptm_df <- ptm_df %>% filter(ptm == i)
ptm_df$FeatureType <- "PTM"
ptm_df <- ptm_df %>% arrange(condition, subset, rowname)

ptm_estimate <- ptm_df %>% select(c("primary", "colname", "condition", "subset")) %>%
                           mutate(rowname = paste(ptm_df$rowname, "estimate"),
                                  ptm = paste(ptm_df$ptm, "estimate"),
                                  Protein = paste(ptm_df$Protein, "estimate"),
                                  FeatureType = "PTM_estimated",
                                  value = NA,
                                  assay = "model")
fixeffects <- rowData(pe[["ptmRel"]])$msqrobModels[[unique(ptm_df$ptm)]] %>% getCoef
ptm_estimate[ptm_estimate$condition=="B"&(ptm_estimate$subset=="x"),]$value <- 
                                                    fixeffects[["(Intercept)"]]
ptm_estimate[ptm_estimate$condition=="B"&(ptm_estimate$subset=="y"),]$value <- 
   fixeffects[["(Intercept)"]] + fixeffects[["subsety"]]
ptm_estimate[ptm_estimate$condition=="A"&(ptm_estimate$subset=="x"),]$value <- 
   fixeffects[["(Intercept)"]] + fixeffects[["conditionA"]]
ptm_estimate[ptm_estimate$condition=="A"&(ptm_estimate$subset=="y"),]$value <- 
   fixeffects[["(Intercept)"]] + fixeffects[["conditionA"]] + fixeffects[["conditionA:subsety"]] + fixeffects[["subsety"]]

plot_df <- rbindlist(list(pepform_df, ptm_df, ptm_estimate), fill = TRUE)
plot_points <- rbindlist(list(pepform_df, ptm_df), fill = TRUE)

plot_df$primary <- forcats::fct_inorder(plot_df$primary)

plots[[i]] <- plot_df %>% ggplot() +
  geom_line(aes(x = primary, y = value , group = rowname, color = FeatureType), size = 2) +
  geom_point(data = plot_points, aes(x = primary, y = value , group = rowname, 
                                     color = FeatureType), size = 5) +
  geom_vline(data=data.frame(x = c(43.5, 17.5, 56.5)),
             aes(xintercept=as.numeric(x)), linetype = "dashed") +
  scale_colour_manual(values = c("Peptide" = "#C3C3C3", "PTM" = "palegreen3",
                                 "PTM_estimated" = "slateblue3")) +
  labs(title = i, x = "Sample", y = "Intensity") +
  theme_light() +
  theme(axis.text.x = element_text(angle = 60, hjust=1, size = 10),
        axis.text.y = element_text(size = 16),
        legend.text=element_text(size=19),
        axis.title.y = element_text(size = 16),
        axis.title.x = element_text(size = 16),
        title = element_text(size = 19),
        strip.text = element_text(size = 16),
        legend.title =  element_blank(),
        legend.direction = "horizontal",
        legend.position = c(0.5, 0.1)) +
  annotate("text", x = 24, y = 6.5, label = "B", size = 8) +
  annotate("text", x = 70, y = 6.5, label = "A", size = 8) +
  annotate("text", x = 9, y = 6, label = "x", size = 7) +
  annotate("text", x = 30, y = 6, label = "y", size = 7) +
  annotate("text", x = 49.5, y = 6, label = "x", size = 7) +
  annotate("text", x = 73, y = 6, label = "y", size = 7) +
  ylim(-6, 6.5)

}
plots
```


### Peptidoform level



```{r, fig.width=18, fig.height=10}
pepf_list1 <- c()
for (contrast in contrasts){
sigTable <- rowData(pe[["pepformRel"]])[[contrast]] %>% filter(adjPval<0.05)
pepf_list1 <- c(pepf_list1, rownames(sigTable))
}
pepf_list1 <- unique(pepf_list1)

plots <- list()
for (i in pepf_list1){
prot_ = rowData(pe[["pepformRel"]])[i,][["Leading.razor.protein"]]
site_ = rowData(pe[["pepformRel"]])[i,][["mod"]]
sites_ = strsplit(site_, ", ")[[1]]
for (site_ in sites_){
ptm_ = paste(prot_, site_, collapse = " ")
print(ptm_)

pepform_df <- longFormat(pe[,,"pepformRel"], rowvars = c("Leading.razor.protein", "ptm"), colvars = c("condition", "subset")) %>% as.data.frame()
pepform_df <- pepform_df %>% filter(Leading.razor.protein == prot_)
pepform_df <- pepform_df %>% filter(grepl(site_, ptm, fixed = T))
pepform_df$FeatureType <- "Peptide"
pepform_df[pepform_df$rowname==i,]$FeatureType <- "SignificantPeptide"
pepform_df <- pepform_df %>% arrange(condition, subset, rowname)

ptm_df <- longFormat(pe[,,"ptmRel"], rowvars = c("protein", "ptm"), colvars = c("condition", "subset")) %>% as.data.frame()
ptm_df <- ptm_df %>% filter(ptm == ptm_)
ptm_df$FeatureType <- "PTM"
ptm_df <- ptm_df %>% arrange(condition, subset, rowname)

ptm_estimate <- ptm_df %>% select(c("primary", "colname", "condition", "subset")) %>%
                           mutate(rowname = paste(ptm_df$rowname, "estimate"),
                                  ptm = paste(ptm_df$ptm, "estimate"),
                                  Protein = paste(ptm_df$Protein, "estimate"),
                                  FeatureType = "PTM_estimated",
                                  value = NA,
                                  assay = "model")
fixeffects <- rowData(pe[["ptmRel"]])$msqrobModels[[unique(ptm_df$ptm)]] %>% getCoef
ptm_estimate[ptm_estimate$condition=="B"&(ptm_estimate$subset=="x"),]$value <- 
                                                    fixeffects[["(Intercept)"]]
ptm_estimate[ptm_estimate$condition=="B"&(ptm_estimate$subset=="y"),]$value <- 
   fixeffects[["(Intercept)"]] + fixeffects[["subsety"]]
ptm_estimate[ptm_estimate$condition=="A"&(ptm_estimate$subset=="x"),]$value <- 
   fixeffects[["(Intercept)"]] + fixeffects[["conditionA"]]
ptm_estimate[ptm_estimate$condition=="A"&(ptm_estimate$subset=="y"),]$value <- 
   fixeffects[["(Intercept)"]] + fixeffects[["conditionA"]] + fixeffects[["conditionA:subsety"]]+ fixeffects[["subsety"]]


plot_df <- rbindlist(list(pepform_df, ptm_df, ptm_estimate), fill = TRUE)
plot_points <- rbindlist(list(pepform_df, ptm_df), fill = TRUE)

# plot_df[plot_df$FeatureType == 'PTM'][['FeatureType']] <- "PTM Summarized"
# plot_df[plot_df$FeatureType != 'PTM'][['FeatureType']] <- "PTM Feature"
plot_df$primary <- forcats::fct_inorder(plot_df$primary)

plots[[paste(i, site_)]] <- plot_df %>% ggplot() +
  geom_line(aes(x = primary, y = value , group = rowname, color = FeatureType), size =2) +
  geom_point(data = plot_points, aes(x = primary, y = value , group = rowname, color = FeatureType), size = 5) +
  geom_vline(data=data.frame(x = c(43.5, 17.5, 56.5)),
             aes(xintercept=as.numeric(x)), linetype = "dashed") +
  scale_colour_manual(values = c("Peptide" = "#C3C3C3", "PTM" = "palegreen3", 
                                 "SignificantPeptide" = "violetred1",
                                 "PTM_estimated" = "lightgoldenrod")) +
  labs(title = paste(i, ptm_), x = "Sample", y = "Intensity") +
  theme_light() +
  theme(axis.text.x = element_text(angle = 60, hjust=1, size = 10),
        axis.text.y = element_text(size = 16),
        legend.text=element_text(size=19),
        axis.title.y = element_text(size = 16),
        axis.title.x = element_text(size = 16),
        title = element_text(size = 19),
        strip.text = element_text(size = 16),
        legend.title =  element_blank(),
        legend.direction = "horizontal",
        legend.position = c(0.5, 0.1)) +
  annotate("text", x = 24, y = 6.5, label = "B", size = 8) +
  annotate("text", x = 70, y = 6.5, label = "A", size = 8) +
  annotate("text", x = 9, y = 6, label = "x", size = 7) +
  annotate("text", x = 30, y = 6, label = "y", size = 7) +
  annotate("text", x = 49.5, y = 6, label = "x", size = 7) +
  annotate("text", x = 73, y = 6, label = "y", size = 7) +
  ylim(-6, 6.5)
}
}
plots
```

