---
title: "simulation visualisations"
output:
  html_document:
    code_download: yes
    theme: cosmo
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
    highlight: tango
    number_sections: yes
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: '4'
---

```{r}
library(readxl)
library(tidyverse)
```

PTM LEVEL

```{r, setup, include=F}
knitr::opts_knit$set(root.dir="C:/Users/Nina/OneDrive - UGent/Documenten/Doctoraat/msqrobPTM paper/simulations/")
```


```{r}
load("MSstats_models/adjusted_models_sim1.rda")
load("MSstats_models/adjusted_models_sim2.rda")
```

```{r}
sign_table11 <- readRDS("simulation 1/sign_table11.rds")
sign_table12 <- readRDS("simulation 1/sign_table12.rds")
sign_table13 <- readRDS("simulation 1/sign_table13.rds")

sign_table21 <- readRDS("simulation 2/sign_table21.rds")
sign_table22 <- readRDS("simulation 2/sign_table22.rds")
sign_table23 <- readRDS("simulation 2/sign_table23.rds")
```

```{r}
msqrob_simulation1 <- rbind(sign_table11, sign_table12, sign_table13)
msqrob_simulation2 <- rbind(sign_table21, sign_table22, sign_table23)
```


```{r}
s <- c(.2,.3)
reps <- c(2,3,5,10)
cond <- c(2,3,4)
param_combos <- expand.grid(s, reps, cond)
```


## msstats type visualisations

Some of the functions below were found in the simulation analysis files on Massive from the MSstats team

## sensitivy - fdp plots
### MSstats

Function to calculate TPR and FDP

```{r}
tprFdp <- function(df, pval, tp){
  ord <- order(df[[pval]])
  df <- df[ord,]
  df$tpr = as.numeric(cumsum(df[[tp]])/sum(df[[tp]]))
  df$fdp = as.numeric(cumsum(!df[[tp]])/(1:length(df[[tp]])))
  df$fpr = as.numeric(cumsum(!df[[tp]])/sum(!df[[tp]]))
  df
}
```


#### Simulation 1

```{r}
for (i in 1:length(adjusted_models_sim1)){
  adjusted_models_sim1[[i]] <- adjusted_models_sim1[[i]] %>%
                                mutate(sd = param_combos[i,1], reps = param_combos[i,2],
                                       conditions = param_combos[i,3], model= "MSstats",
                                       change = !grepl("NoChange", adjusted_models_sim1[[i]]$GlobalProtein))
}
```

```{r}
for (i in 1:length(adjusted_models_sim1)){
  adjusted_models_sim1[[i]] <- tprFdp(adjusted_models_sim1[[i]], "pvalue", "change")
}
```

```{r}
point_data <- tibble()
for (i in 1:24){
  df <- adjusted_models_sim1[[i]]
  row_ = sum(df$adj.pvalue < 0.05, na.rm =T)
  if (row_ == 0){
    point = tibble(x = 0, y = 0, fpr = 0,
                   sd = as.factor(df[1,]$sd), 
                   reps = as.factor(df[1,]$reps), 
                   conditions = as.factor(df[1,]$conditions),
                   model = "MSstats")
  }
  else {
  x = df[row_,]$fdp
  fpr = df[row_,]$fpr
  y = df[row_,]$tpr
  sd = as.factor(df[row_,]$sd)
  reps = as.factor(df[row_,]$reps)
  conditions = as.factor(df[row_,]$conditions)
  point = tibble(x = x, y = y, fpr = fpr, sd = sd, reps = reps, conditions = conditions, model = "MSstats")
  }
  point_data = rbind(point_data, point)
}
```

```{r, fig.height=7, fig.width=12}
model_df <- do.call("rbind", adjusted_models_sim1)

model_df$reps = as.factor(model_df$reps)
model_df %>%
      ggplot(aes(x = fdp, y = tpr, color = reps)) +
      geom_path() +
      geom_vline(xintercept=0.05,lty=2) +
      geom_point(data = point_data, aes(x, y, color = reps)) +
      facet_grid(vars(sd), vars(conditions))
```

ROC curve

```{r, fig.height=7, fig.width=12}
model_df <- do.call("rbind", adjusted_models_sim1)

model_df$reps = as.factor(model_df$reps)
model_df %>%
      ggplot(aes(x = fpr, y = tpr, color = reps)) +
      geom_path() +
      geom_vline(xintercept=0.05,lty=2) +
      geom_point(data = point_data, aes(fpr, y, color = reps)) +
      facet_grid(vars(sd), vars(conditions))
```

```{r}
for (i in 1:24){
  df <- adjusted_models_sim1[[i]]
  p <- df %>% ggplot(aes(x = fdp, y = tpr)) +
      geom_path() +
      geom_vline(xintercept=0.01,lty=2) +
      geom_point(data = adjusted_models_sim1[[i]][sum(adjusted_models_sim1[[i]]$adj.pvalue < 0.05, na.rm = TRUE), ],
             aes(x = fdp, y = tpr), cex = 2, color = "blue", size = 3) +
          ggtitle(paste("sd:", as.character(param_combos[i, 1]), "reps:", 
                    as.character(param_combos[i,2]), "conditions:", 
                    as.character(param_combos[i,3])))
  print(p)
}
```



#### Simulation 2

```{r}
for (i in 1:length(adjusted_models_sim2)){
  adjusted_models_sim2[[i]] <- adjusted_models_sim2[[i]] %>%
                                mutate(sd = param_combos[i,1], reps = param_combos[i,2],
                                       conditions = param_combos[i,3], model= "MSstats",
                                       change = !grepl("NoChange", adjusted_models_sim2[[i]]$GlobalProtein))
}
```

```{r}
for (i in 1:length(adjusted_models_sim2)){
  adjusted_models_sim2[[i]] <- tprFdp(adjusted_models_sim2[[i]], "pvalue", "change")
}
```

```{r}
point_data_2 <- tibble()
for (i in 1:24){
  df <- adjusted_models_sim2[[i]]
  row_ = sum(df$adj.pvalue < 0.05, na.rm =T)
  if (row_ == 0){
    point = tibble(x = 0, y = 0, fpr = 0,
                   sd = as.factor(df[1,]$sd), 
                   reps = as.factor(df[1,]$reps), 
                   conditions = as.factor(df[1,]$conditions),
                   model = "MSstats")
  }
  else {
  x = df[row_,]$fdp
  fpr = df[row_,]$fpr
  y = df[row_,]$tpr
  sd = as.factor(df[row_,]$sd)
  reps = as.factor(df[row_,]$reps)
  conditions = as.factor(df[row_,]$conditions)
  point = tibble(x = x, y = y, fpr = fpr, sd = sd, reps = reps, conditions = conditions, model = "MSstats")
  }
  point_data_2 = rbind(point_data_2, point)
}
```

```{r, fig.height=7, fig.width=12}
model_df2 <- do.call("rbind", adjusted_models_sim2)
model_df2$reps = as.factor(model_df2$reps)

model_df2 %>%
      ggplot(aes(x = fdp, y = tpr, color = reps)) +
      geom_path() +
      geom_vline(xintercept=0.05,lty=2) +
      geom_point(data = point_data_2, aes(x, y, color = reps)) +
      facet_grid(vars(sd), vars(conditions))
```

ROC curve

```{r, fig.height=7, fig.width=12}
model_df2 %>%
      ggplot(aes(x = fpr, y = tpr, color = reps)) +
      geom_path() +
      geom_vline(xintercept=0.05,lty=2) +
      geom_point(data = point_data_2, aes(fpr, y, color = reps)) +
      facet_grid(vars(sd), vars(conditions))
```

```{r}
for (i in 1:24){
  df <- adjusted_models_sim2[[i]]
  p <- df %>% ggplot(aes(x = fdp, y = tpr)) +
      geom_path() +
      geom_vline(xintercept=0.01,lty=2) +
      geom_point(data = adjusted_models_sim2[[i]][sum(adjusted_models_sim2[[i]]$adj.pvalue < 0.05, na.rm = TRUE), ],
             aes(x = fdp, y = tpr), cex = 2, color = "blue", size = 3) +
      ggtitle(paste("sd:", as.character(param_combos[i, 1]), "reps:", 
                    as.character(param_combos[i,2]), "conditions:", 
                    as.character(param_combos[i,3])))
  print(p)
}
```

### msqrob

#### Simulation 1

```{r}
msqrob_simulation1 <- 
  msqrob_simulation1 %>%     
  rownames_to_column("Protein") %>%
  mutate(change = !grepl("NoChange", Protein),
         dataset = paste(sd, reps, conditions, sep = "_"),
         model = "msqrob"
)
```

```{r}
msqrob_simulation1$tpr <- 1000
msqrob_simulation1$fdp <- 1000
msqrob_simulation1$fpr <- 1000
for (i in unique(msqrob_simulation1$dataset)){
  msqrob_simulation1[msqrob_simulation1$dataset==i,] <- tprFdp(msqrob_simulation1[msqrob_simulation1$dataset==i,], "pval", "change")
}
```

```{r}
point_data_3 <- tibble()
for (i in unique(msqrob_simulation1$dataset)){
  df <- msqrob_simulation1[msqrob_simulation1$dataset==i,]
  row_ = sum(df$adjPval < 0.05, na.rm =T)
  if (row_ == 0){
    point = tibble(x = 0, y = 0, fpr = 0,
                   sd = as.factor(df[1,]$sd), 
                   reps = as.factor(df[1,]$reps), 
                   conditions = as.factor(df[1,]$conditions))
  }
  else {
  x = df[row_,]$fdp
  fpr = df[row_,]$fpr
  y = df[row_,]$tpr
  sd = as.factor(df[row_,]$sd)
  reps = as.factor(df[row_,]$reps)
  conditions = as.factor(df[row_,]$conditions)
  point = tibble(x = x, y = y, fpr = fpr, sd = sd, reps = reps, conditions = conditions, model = "msqrob")
  }
  point_data_3 = rbind(point_data_3, point)
}
```

```{r, fig.height=10, fig.width=11}
msqrob_simulation1$reps = as.factor(msqrob_simulation1$reps)

msqrob_simulation1 %>%
      ggplot(aes(x = fdp, y = tpr, color = reps)) +
      geom_path(size=0.8) +
      geom_vline(xintercept=0.05,lty=2) +
      geom_point(data = point_data_3, aes(x, y, color = reps), size=2) +
      facet_grid(vars(conditions), vars(sd), labeller = label_both) +
      theme(axis.text=element_text(size=11),
            axis.title=element_text(size=15),
            strip.text.x = element_text(size = 18),
            strip.text.y = element_text(size = 18),
            legend.title = element_text(size=14), 
            legend.text = element_text(size=11))
```

ROC curve

```{r, fig.height=10, fig.width=11}
msqrob_simulation1$reps = as.factor(msqrob_simulation1$reps)

msqrob_simulation1 %>%
      ggplot(aes(x = fpr, y = tpr, color = reps)) +
      geom_path(size=0.8) +
      geom_vline(xintercept=0.05,lty=2) +
      geom_point(data = point_data_3, aes(fpr, y, color = reps), size=2) +
      facet_grid(vars(conditions), vars(sd), labeller = label_both) +
      theme(axis.text=element_text(size=11),
            axis.title=element_text(size=15),
            strip.text.x = element_text(size = 18),
            strip.text.y = element_text(size = 18),
            legend.title = element_text(size=14), 
            legend.text = element_text(size=11))
```

```{r}
j <- 1
for (i in unique(msqrob_simulation1$dataset)){
  df <- msqrob_simulation1[msqrob_simulation1$dataset==i,]
  p <- df %>% ggplot(aes(x = fdp, y = tpr)) +
      geom_path() +
      geom_vline(xintercept=0.01,lty=2) +
      geom_point(data = df[sum(df$adjPval < 0.05, na.rm = TRUE), ],
             aes(x = fdp, y = tpr), cex = 2, color = "blue", size = 3) +
          ggtitle(paste("sd:", as.character(param_combos[j, 1]), "reps:", 
                    as.character(param_combos[j,2]), "conditions:", 
                    as.character(param_combos[j,3])))
  print(p)
  j <- j + 1
}
```

#### Simulation 2

```{r}
msqrob_simulation2 <- 
  msqrob_simulation2 %>%     
  rownames_to_column("Protein") %>%
  mutate(change = !grepl("NoChange", Protein),
         dataset = paste(sd, reps, conditions, sep = "_"),
         model = "msqrob"
)
```

```{r}
msqrob_simulation2$tpr <- 1000
msqrob_simulation2$fdp <- 1000
msqrob_simulation2$fpr <- 1000
for (i in unique(msqrob_simulation2$dataset)){
  msqrob_simulation2[msqrob_simulation2$dataset==i,] <- tprFdp(msqrob_simulation2[msqrob_simulation2$dataset==i,], "pval", "change")
}
```

```{r}
point_data_4 <- tibble()
for (i in unique(msqrob_simulation2$dataset)){
  df <- msqrob_simulation2[msqrob_simulation2$dataset==i,]
  row_ = sum(df$adjPval < 0.05, na.rm =T)
  if (row_ == 0){
    point = tibble(x = 0, y = 0, fpr = 0,
                   sd = as.factor(df[1,]$sd), 
                   reps = as.factor(df[1,]$reps), 
                   conditions = as.factor(df[1,]$conditions))
  }
  else {
  x = df[row_,]$fdp
  fpr = df[row_,]$fpr
  y = df[row_,]$tpr
  sd = as.factor(df[row_,]$sd)
  reps = as.factor(df[row_,]$reps)
  conditions = as.factor(df[row_,]$conditions)
  point = tibble(x = x, y = y, fpr = fpr, sd = sd, reps = reps, conditions = conditions, model = "msqrob")
  }
  point_data_4 = rbind(point_data_4, point)
}
```

```{r, fig.height=10, fig.width=11}
msqrob_simulation2$reps = as.factor(msqrob_simulation2$reps)

msqrob_simulation2 %>%
      ggplot(aes(x = fdp, y = tpr, color = reps)) +
      geom_path(size=0.8) +
      geom_vline(xintercept=0.05,lty=2) +
      geom_point(data = point_data_4, aes(x, y, color = reps), size=2) +
      facet_grid(vars(conditions), vars(sd), labeller = label_both) +
      theme(axis.text=element_text(size=11),
      axis.title=element_text(size=15),
      strip.text.x = element_text(size = 18),
      strip.text.y = element_text(size = 18),
      legend.title = element_text(size=14), 
      legend.text = element_text(size=11))
```

ROC curve

```{r, fig.height=10, fig.width=11}
msqrob_simulation2$reps = as.factor(msqrob_simulation2$reps)

msqrob_simulation2 %>%
      ggplot(aes(x = fpr, y = tpr, color = reps)) +
      geom_path(size=0.8) +
      geom_vline(xintercept=0.05,lty=2) +
      geom_point(data = point_data_4, aes(fpr, y, color = reps), size=2) +
      facet_grid(vars(conditions), vars(sd), labeller = label_both) +
      theme(axis.text=element_text(size=11),
      axis.title=element_text(size=15),
      strip.text.x = element_text(size = 18),
      strip.text.y = element_text(size = 18),
      legend.title = element_text(size=14), 
      legend.text = element_text(size=11))
```

```{r}
j <- 1
for (i in unique(msqrob_simulation2$dataset)){
  df <- msqrob_simulation2[msqrob_simulation2$dataset==i,]
  p <- df %>% ggplot(aes(x = fdp, y = tpr)) +
      geom_path() +
      geom_vline(xintercept=0.01,lty=2) +
      geom_point(data = df[sum(df$adjPval < 0.05, na.rm = TRUE), ],
             aes(x = fdp, y = tpr), cex = 2, color = "blue", size = 3) +
          ggtitle(paste("sd:", as.character(param_combos[j, 1]), "reps:", 
                    as.character(param_combos[j,2]), "conditions:", 
                    as.character(param_combos[j,3])))
  print(p)
  j <- j + 1
}
```

### together

#### Simulation 1

```{r}
model_df <- model_df %>%
              dplyr::rename(logFC = log2FC,
                     se = SE,
                     t = Tvalue,
                     df = DF,
                     pval = pvalue,
                     adjPval = adj.pvalue) %>% 
              mutate(dataset = paste(sd, reps, conditions, sep = "_")) %>%
              select(-GlobalProtein)

model_df_together <- rbind(model_df, msqrob_simulation1)
```

```{r}
point_data_together <- rbind(point_data, point_data_3)
```

```{r}
get_point_data <- function(df, adjpval, alpha, model){
  row_ = sum(df[[adjpval]] < alpha, na.rm =T)
  if (row_ == 0){
    point = tibble(x = 0, y = 0, fpr = 0,
                   sd = as.factor(df[1,]$sd), 
                   reps = as.factor(df[1,]$reps), 
                   conditions = as.factor(df[1,]$conditions),
                   alpha = alpha,
                   model = model)
  }
  else {
    point= tibble(x = df[row_,]$fdp,
                  fpr = df[row_,]$fpr,
                  y = df[row_,]$tpr,
                  sd = as.factor(df[row_,]$sd),
                  reps = as.factor(df[row_,]$reps),
                  conditions = as.factor(df[row_,]$conditions),
                  alpha = alpha,
                  model = model)
  }
  return(point)
}

```

```{r}
point_data_sim1_msstats <- lapply(seq(1:24), function(x) {
  p1 = get_point_data(adjusted_models_sim1[[x]], "adj.pvalue", 0.01, "MSstats")
  p5 = get_point_data(adjusted_models_sim1[[x]], "adj.pvalue", 0.05, "MSstats")
  p10 = get_point_data(adjusted_models_sim1[[x]], "adj.pvalue", 0.1, "MSstats")
  return(rbind(p1, p5, p10))
})
point_data_sim1_msstats = do.call("rbind", point_data_sim1_msstats)

point_data_sim1_msqrob <- lapply(unique(msqrob_simulation1$dataset), function(x) {
  p1 = get_point_data(msqrob_simulation1[msqrob_simulation1$dataset==x,], "adjPval", 0.01, "msqrob")
  p5 = get_point_data(msqrob_simulation1[msqrob_simulation1$dataset==x,], "adjPval", 0.05, "msqrob")
  p10 = get_point_data(msqrob_simulation1[msqrob_simulation1$dataset==x,], "adjPval", 0.1, "msqrob")
  return(rbind(p1, p5, p10))
})
point_data_sim1_msqrob = do.call("rbind", point_data_sim1_msqrob)
point_data_sim1_together = rbind(point_data_sim1_msstats, point_data_sim1_msqrob)
point_data_sim1_together$alpha <- as.factor(point_data_sim1_together$alpha)
```


```{r, fig.height=10, fig.width=13}
model_df_together %>%
      ggplot(aes(x = fdp, y = tpr, color = reps)) +
      geom_path(aes(linetype = model), size = 1) +
      geom_vline(xintercept=0.05,lty=2) +
      geom_point(data = point_data_sim1_together, aes(x = x, y = y, shape = alpha, fill = reps), size = 1.5, color = "black") +
      facet_grid(vars(sd), vars(conditions), labeller = label_both)
```

Same plots, but more aesthetically pleasing (according to me, anyway)

```{r, fig.height=10, fig.width=11}
model_df_together %>%
      ggplot(aes(x = fdp, y = tpr, color = reps)) +
      geom_path(aes(linetype = model, size = model)) +
      scale_size_manual(values = c(msqrob = 1.15, MSstats = 0.8)) +
      scale_linetype_manual(values = c(msqrob = "solid", MSstats = "dotted")) +
      geom_vline(xintercept=0.05,lty=2) +
      geom_point(data = point_data_together, 
                 aes(x = x, y = y, shape = model, color = reps), size = 2.5) +
      facet_grid(vars(conditions), vars(sd), labeller = label_both) +
      theme(axis.text=element_text(size=11),
      axis.title=element_text(size=15),
      strip.text.x = element_text(size = 18),
      strip.text.y = element_text(size = 18),
      legend.title = element_text(size=14), 
      legend.text = element_text(size=11))
```

ROC curves

```{r, fig.height=10, fig.width=11}
model_df_together %>%
      ggplot(aes(x = fpr, y = tpr, color = reps)) +
      geom_path(aes(linetype = model, size = model)) +
      scale_size_manual(values = c(msqrob = 1.1, MSstats = 1.1)) +
      scale_linetype_manual(values = c(msqrob = "solid", MSstats = "dotted")) +
      geom_vline(xintercept=0.05,lty=2) +
      #geom_point(data = point_data_together, aes(x = fpr, y = y, shape = model), size = 2, color = "gray30") +
      facet_grid(vars(conditions), vars(sd), labeller = label_both) +
      theme(axis.text=element_text(size=11),
      axis.title=element_text(size=15),
      strip.text.x = element_text(size = 18),
      strip.text.y = element_text(size = 18),
      legend.title = element_text(size=14), 
      legend.text = element_text(size=11))
```

#### Simulation 2

```{r}
model_df2 <- model_df2 %>%
              dplyr::rename(logFC = log2FC,
                     se = SE,
                     t = Tvalue,
                     df = DF,
                     pval = pvalue,
                     adjPval = adj.pvalue) %>% 
              mutate(dataset = paste(sd, reps, conditions, sep = "_")) %>%
              select(-GlobalProtein)

model_df2_together <- rbind(model_df2, msqrob_simulation2)
```

```{r}
point_data_together_2 <- rbind(point_data_2, point_data_4)
```


```{r}
point_data_sim2_msstats <- lapply(seq(1:24), function(x) {
  p1 = get_point_data(adjusted_models_sim2[[x]], "adj.pvalue", 0.01, "MSstats")
  p5 = get_point_data(adjusted_models_sim2[[x]], "adj.pvalue", 0.05, "MSstats")
  p10 = get_point_data(adjusted_models_sim2[[x]], "adj.pvalue", 0.1, "MSstats")
  return(rbind(p1, p5, p10))
})
point_data_sim2_msstats = do.call("rbind", point_data_sim2_msstats)

point_data_sim2_msqrob <- lapply(unique(msqrob_simulation2$dataset), function(x) {
  p1 = get_point_data(msqrob_simulation2[msqrob_simulation2$dataset==x,], "adjPval", 0.01, "msqrob")
  p5 = get_point_data(msqrob_simulation2[msqrob_simulation2$dataset==x,], "adjPval", 0.05, "msqrob")
  p10 = get_point_data(msqrob_simulation2[msqrob_simulation2$dataset==x,], "adjPval", 0.1, "msqrob")
  return(rbind(p1, p5, p10))
})
point_data_sim2_msqrob = do.call("rbind", point_data_sim2_msqrob)
point_data_sim2_together = rbind(point_data_sim2_msstats, point_data_sim2_msqrob)
point_data_sim2_together$alpha <- as.factor(point_data_sim2_together$alpha)
```


```{r, fig.height=10, fig.width=13}
model_df2_together %>%
      ggplot(aes(x = fdp, y = tpr, color = reps)) +
      geom_path(aes(linetype = model), size = 1) +
      geom_vline(xintercept=0.05,lty=2) +
      geom_point(data = point_data_sim2_together, aes(x = x, y = y, shape = alpha, fill = reps), size = 1.5, color = "black") +
      facet_grid(vars(sd), vars(conditions), labeller = label_both)
```


```{r, fig.height=10, fig.width=11}
model_df2_together %>%
      ggplot(aes(x = fdp, y = tpr, color = reps)) +
      geom_path(aes(linetype = model, size = model)) +
      scale_size_manual(values = c(msqrob = 1.15, MSstats = 0.8)) +
      scale_linetype_manual(values = c(msqrob = "solid", MSstats = "dotted")) +
      geom_vline(xintercept=0.05,lty=2) +
      geom_point(data = point_data_together_2, aes(x = x, y = y, shape = model, color = reps), size = 2.5) +
      facet_grid(vars(conditions), vars(sd), labeller = label_both) +
      theme(axis.text=element_text(size=11),
      axis.title=element_text(size=15),
      strip.text.x = element_text(size = 18),
      strip.text.y = element_text(size = 18),
      legend.title = element_text(size=14), 
      legend.text = element_text(size=11))
```

ROC curves

```{r, fig.height=10, fig.width=11}
model_df2_together %>%
      ggplot(aes(x = fpr, y = tpr, color = reps)) +
      geom_path(aes(linetype = model, size = model)) +
      scale_size_manual(values = c(msqrob = 1.1, MSstats = 1.1)) +
      scale_linetype_manual(values = c(msqrob = "solid", MSstats = "dotted")) +
      geom_vline(xintercept=0.05,lty=2) +
      #geom_point(data = point_data_together_2, aes(x = fpr, y = y, shape = model), size = 2, color = "gray35") +
      facet_grid(vars(conditions), vars(sd), labeller = label_both) +
      theme(axis.text=element_text(size=11),
      axis.title=element_text(size=15),
      strip.text.x = element_text(size = 18),
      strip.text.y = element_text(size = 18),
      legend.title = element_text(size=14), 
      legend.text = element_text(size=11))
```



## p value distributions

### msstats

#### Simulation 1

```{r}
pvals <- c()
for (i in 1:length(adjusted_models_sim1)){
  pvals <- c(pvals, df %>%
    filter(change == F) %>%
    pull(pval))

}
hist(pvals)
```

```{r}
pvals <- c()
for (i in 1:length(adjusted_models_sim1)){
  pvals <- c(pvals, df %>%
    filter(change == T) %>%
    pull(pval))

}
hist(pvals)
```


### msqrob

#### Simulation 1

```{r}
pvals <- c()
for (i in unique(msqrob_simulation1$dataset)){
  df <- msqrob_simulation1[msqrob_simulation1$dataset==i,]
  pvals <- c(pvals, df %>%
    filter(grepl(pattern = "NoChange", Protein)) %>%
      pull(pval))
}

hist(pvals)
```

```{r}
pvals <- c()
for (i in unique(msqrob_simulation1$dataset)){
  df <- msqrob_simulation1[msqrob_simulation1$dataset==i,]
  pvals <- c(pvals, df %>%
    filter(!grepl(pattern = "NoChange", Protein)) %>%
      pull(pval))
}

hist(pvals)
```


#### Simulation 2

```{r}
pvals <- c()
for (i in unique(msqrob_simulation2$dataset)){
  df <- msqrob_simulation2[msqrob_simulation2$dataset==i,]
  pvals <- c(pvals, df %>%
    filter(grepl(pattern = "NoChange", Protein)) %>%
      pull(pval))
}

hist(pvals)
```

```{r}
pvals <- c()
for (i in unique(msqrob_simulation2$dataset)){
  df <- msqrob_simulation2[msqrob_simulation2$dataset==i,]
  pvals <- c(pvals, df %>%
    filter(!grepl(pattern = "NoChange", Protein)) %>%
      pull(pval))
}

hist(pvals)
```



### Versus when starting from summarised data

```{r}
load("C:/Users/Nina/OneDrive - UGent/Documenten/Doctoraat/msqrobPTM/Datasets MSStats/simulation_data/signtable_simulation1_msqrob_comparison1_startfromsummary.rda")
sign_table11 <- sign_table
load("C:/Users/Nina/OneDrive - UGent/Documenten/Doctoraat/msqrobPTM/Datasets MSStats/simulation_data/signtable_simulation1_msqrob_comparison2_startfromsummary.rda")
sign_table12 <- sign_table
load("C:/Users/Nina/OneDrive - UGent/Documenten/Doctoraat/msqrobPTM/Datasets MSStats/simulation_data/signtable_simulation1_msqrob_comparison3_startfromsummary.rda")
sign_table13 <- sign_table

load("C:/Users/Nina/OneDrive - UGent/Documenten/Doctoraat/msqrobPTM/Datasets MSStats/simulation_data/signtable_simulation2_msqrob_comparison1_startfromsummary.rda")
sign_table21 <- sign_table
load("C:/Users/Nina/OneDrive - UGent/Documenten/Doctoraat/msqrobPTM/Datasets MSStats/simulation_data/signtable_simulation2_msqrob_comparison2_startfromsummary.rda")
sign_table22 <- sign_table
load("C:/Users/Nina/OneDrive - UGent/Documenten/Doctoraat/msqrobPTM/Datasets MSStats/simulation_data/signtable_simulation2_msqrob_comparison3_startfromsummary.rda")
sign_table23 <- sign_table
```

```{r}
msqrob_simulation1_summary <- rbind(sign_table11, sign_table12, sign_table13)
msqrob_simulation2_summary <- rbind(sign_table21, sign_table22, sign_table23)
```

#### Simulation 1

```{r}
point_data_7 <- tibble()
for (i in unique(msqrob_simulation1_summary$dataset)){
  df <- msqrob_simulation1_summary[msqrob_simulation1_summary$dataset==i,]
  row_ = sum(df$adjPval < 0.05, na.rm =T)
  if (row_ == 0){
    point = tibble(x = 0, y = 0, fpr = 0,
                   sd = as.factor(df[1,]$sd), 
                   reps = as.factor(df[1,]$reps), 
                   conditions = as.factor(df[1,]$conditions))
  }
  else {
  x = df[row_,]$fdp
  fpr = df[row_,]$fpr
  y = df[row_,]$tpr
  sd = as.factor(df[row_,]$sd)
  reps = as.factor(df[row_,]$reps)
  conditions = as.factor(df[row_,]$conditions)
  point = tibble(x = x, y = y, fpr = fpr, sd = sd, reps = reps, conditions = conditions, model = "msqrob_summaryStart")
  }
  point_data_7 = rbind(point_data_7, point)
}
```

```{r}
msqrob_simulation1_summary <- 
  msqrob_simulation1_summary %>%     
  rownames_to_column("Protein") %>%
  mutate(change = !grepl("NoChange", Protein),
         dataset = paste(sd, reps, conditions, sep = "_"),
         model = "msqrob_summaryStart"
)
```

```{r}
msqrob_simulation1_summary$tpr <- 1000
msqrob_simulation1_summary$fdp <- 1000
msqrob_simulation1_summary$fpr <- 1000
for (i in unique(msqrob_simulation1_summary$dataset)){
  msqrob_simulation1_summary[msqrob_simulation1_summary$dataset==i,] <- 
    tprFdp(msqrob_simulation1_summary[msqrob_simulation1_summary$dataset==i,], "pval", "change")
}
```


```{r}
simulation1_together <- rbind(msqrob_simulation1_summary, msqrob_simulation1)
```

```{r}
point_data_together <- rbind(point_data_3, point_data_7)
```


```{r, fig.height=10, fig.width=13}
simulation1_together %>%
      ggplot(aes(x = fdp, y = tpr, color = reps)) +
      geom_path(aes(linetype = model), size = 1) +
      geom_vline(xintercept=0.05,lty=2) +
      geom_point(data = point_data_together, aes(x = x, y = y, shape = model), size = 1.5, color = "black") +
      facet_grid(vars(sd), vars(conditions), labeller = label_both)
```

```{r, fig.height=10, fig.width=11}
simulation1_together %>%
      ggplot(aes(x = fpr, y = tpr, color = reps)) +
      geom_path(aes(linetype = model, size = model)) +
      scale_size_manual(values = c(msqrob = 1.1, msqrob_summaryStart = 1.1)) +
      scale_linetype_manual(values = c(msqrob = "solid", msqrob_summaryStart = "dotted")) +
      geom_vline(xintercept=0.05,lty=2) +
      #geom_point(data = point_data_together_2, aes(x = fpr, y = y, shape = model), size = 2, color = "gray35") +
      facet_grid(vars(conditions), vars(sd), labeller = label_both) +
      theme(axis.text=element_text(size=11),
      axis.title=element_text(size=15),
      strip.text.x = element_text(size = 18),
      strip.text.y = element_text(size = 18),
      legend.title = element_text(size=14), 
      legend.text = element_text(size=11))
```



#### Simulation 2

```{r}
point_data_8 <- tibble()
for (i in unique(msqrob_simulation2_summary$dataset)){
  df <- msqrob_simulation2_summary[msqrob_simulation2_summary$dataset==i,]
  row_ = sum(df$adjPval < 0.05, na.rm =T)
  if (row_ == 0){
    point = tibble(x = 0, y = 0, fpr = 0,
                   sd = as.factor(df[1,]$sd), 
                   reps = as.factor(df[1,]$reps), 
                   conditions = as.factor(df[1,]$conditions))
  }
  else {
  x = df[row_,]$fdp
  fpr = df[row_,]$fpr
  y = df[row_,]$tpr
  sd = as.factor(df[row_,]$sd)
  reps = as.factor(df[row_,]$reps)
  conditions = as.factor(df[row_,]$conditions)
  point = tibble(x = x, y = y, fpr = fpr, sd = sd, reps = reps, conditions = conditions, model = "msqrob_summaryStart")
  }
  point_data_8 = rbind(point_data_8, point)
}
```

```{r}
msqrob_simulation2_summary <- 
  msqrob_simulation2_summary %>%     
  rownames_to_column("Protein") %>%
  mutate(change = !grepl("NoChange", Protein),
         dataset = paste(sd, reps, conditions, sep = "_"),
         model = "msqrob_summaryStart"
)
```

```{r}
msqrob_simulation2_summary$tpr <- 1000
msqrob_simulation2_summary$fdp <- 1000
msqrob_simulation2_summary$fpr <- 1000
for (i in unique(msqrob_simulation2_summary$dataset)){
  msqrob_simulation2_summary[msqrob_simulation2_summary$dataset==i,] <- 
    tprFdp(msqrob_simulation2_summary[msqrob_simulation2_summary$dataset==i,], "pval", "change")
}
```


```{r}
simulation2_together <- rbind(msqrob_simulation2_summary, msqrob_simulation2)
```

```{r}
point_data_together <- rbind(point_data_4, point_data_8)
```


```{r, fig.height=10, fig.width=13}
simulation2_together %>%
      ggplot(aes(x = fdp, y = tpr, color = reps)) +
      geom_path(aes(linetype = model), size = 1) +
      geom_vline(xintercept=0.05,lty=2) +
      geom_point(data = point_data_together, aes(x = x, y = y, shape = model), size = 1.5, color = "black") +
      facet_grid(vars(sd), vars(conditions), labeller = label_both)
```

```{r, fig.height=10, fig.width=11}
simulation2_together %>%
      ggplot(aes(x = fpr, y = tpr, color = reps)) +
      geom_path(aes(linetype = model, size = model)) +
      scale_size_manual(values = c(msqrob = 1.1, msqrob_summaryStart = 1.1)) +
      scale_linetype_manual(values = c(msqrob = "solid", msqrob_summaryStart = "dotted")) +
      geom_vline(xintercept=0.05,lty=2) +
      #geom_point(data = point_data_together_2, aes(x = fpr, y = y, shape = model), size = 2, color = "gray35") +
      facet_grid(vars(conditions), vars(sd), labeller = label_both) +
      theme(axis.text=element_text(size=11),
      axis.title=element_text(size=15),
      strip.text.x = element_text(size = 18),
      strip.text.y = element_text(size = 18),
      legend.title = element_text(size=14), 
      legend.text = element_text(size=11))
```