---
title: "ubiquitin analysis"
output:
  html_document:
    code_download: yes
    theme: cosmo
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
    highlight: tango
    number_sections: yes
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: '4'
---


```{r}
library(tidyverse)
library(MSstatsPTM)
library(QFeatures)
library(msqrob2)
library(gt)
library(seqinr)
library(data.table)
library(lme4)
```

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir="C:/Users/Nina/OneDrive - UGent/Documenten/Doctoraat/msqrobPTM paper/biological ubiquitin dataset/")
```


```{r}
load("usp30_input_data.rda")
```


``` {r}
#PTM level and Protein level data
head(raw_input_data)
```

# MSstatsPTM analysis

The MSstats workflow and the input data was found on Massive: https://massive.ucsd.edu/ProteoSAFe/dataset_files.jsp?task=f9e6117e00794ce5b2e8802601be16bc#%7B%22table_sort_history%22%3A%22main.collection_asc%22%7D

```{r}
## Convert to MSstatsPTM input format
raw_input_data$FragmentIon <- NA
raw_input_data$ProductCharge <- NA
raw_input_data$IsotopeLabelType <- 'L'
raw_input_data <- raw_input_data %>% 
  mutate(PrecursorCharge = stringr::str_split(
    raw_input_data$feature, "_", simplify = TRUE)[,2])

colnames(raw_input_data) <- c('ProteinName', 'PeptideSequence', 'is_mod', 'Feature', 
                            'Condition', "Batch", "Run", 'Intensity', 'site', 'FragmentIon',
                            'ProductCharge', 'IsotopeLabelType', 'PrecursorCharge')

raw_input_data$BioReplicate <- raw_input_data$Batch
raw_input_data$Intensity <- 2^raw_input_data$Intensity
```



```{r}
## Split modified and unmodified peptides
PTM_df <- raw_input_data %>% filter(is_mod == TRUE)
PTM_df$ProteinName <- paste(PTM_df$ProteinName, PTM_df$site, sep = '_')
protein_df <- raw_input_data %>% filter(is_mod == FALSE)
```


```{r}
## Remove unneeded columns
drops <- c("is_mod", "site", "Batch", "Feature")
PTM_df <- PTM_df[ , !(names(PTM_df) %in% drops)]
protein_df <- protein_df[ , !(names(protein_df) %in% drops)]

input_df <- list('PTM' = PTM_df, 'PROTEIN' = protein_df)
```

## summarisation

```{r}
## Summarize features
summarized_ptm <- dataSummarizationPTM(input_df, verbose = F)
## Model summarized results
## Full pairwise comparison used
model_ptm <- groupComparisonPTM(summarized_ptm, data.type = "LabelFree")
```

```{r eval=FALSE, include=FALSE}
save(summarized_ptm, file ="C:/Users/Nina/OneDrive - UGent/Documenten/Doctoraat/msqrobPTM/biological ubiquitin dataset/MSstatsSummarizedData.RData")
save(model_ptm, file = "C:/Users/Nina/OneDrive - UGent/Documenten/Doctoraat/msqrobPTM/biological ubiquitin dataset/MSstatsModel.rda")
```

# Msqrob2PTM analysis

## Start from raw data

```{r}
load("usp30_input_data.rda")
raw_input_data <- raw_input_data %>% arrange(run)
pepform_df <- pivot_wider(raw_input_data, id_cols = c("protein", "peptide", "feature", "site"),
                           names_from = "run", values_from = "log2inty")
```

```{r}
head(pepform_df, 10)
```

```{r}
pepform_df = pepform_df %>%
                  mutate(site = case_when(site == "None" ~ "",
                        TRUE ~ site)) %>% 
                  mutate(ptm = case_when(
                    site != "" ~ paste(protein, site, sep = "_"),
                        TRUE ~ "")) 
head(pepform_df, 20)
```

### Transform into QFeatures object

```{r}
#order dataframe by protein, for the normalisation step
pepform_df = pepform_df[order(pepform_df$protein),]
pe <- readQFeatures(pepform_df,ecol= 5:20,name="peptidoformRaw", fnames = "feature")
```

```{r}
colData(pe)$group <- sapply(str_split(rownames(colData(pe)), "-"), function(x) x[1])
colData(pe)$biorep <- gsub("T[0-9]", "", rownames(colData(pe)))
colData(pe)$techrep <- rownames(colData(pe))

colData(pe)$group <- relevel(as.factor(colData(pe)$group), ref="Ctrl")
colData(pe)$biorep <- as.factor(colData(pe)$biorep)
colData(pe)$techrep <- as.factor(colData(pe)$techrep)

colData(pe)
```

### Preprocessing

```{r}
print(nrow(pe[["peptidoformRaw"]]))
rowData(pe[["peptidoformRaw"]])$nNonZero <- rowSums(assay(pe[["peptidoformRaw"]]) > 0, na.rm = T)
pe <- zeroIsNA(pe, i = "peptidoformRaw")
pe <- pe[rowData(pe[["peptidoformRaw"]])$nNonZero>2,,]
pe <- normalize(pe, i = "peptidoformRaw", method = "center.median", name = "peptidoform")
print(nrow(pe[["peptidoform"]]))
```

```{r}
boxplot(assay(pe[["peptidoform"]]))
```

### Normalize pepform log intensity by protein log intensity

First one is for visualisation purposes
```{r,warning=FALSE}
pe <- aggregateFeatures(pe,
 i = "peptidoformRaw",
 fcol = "protein",
 na.rm = TRUE,
 name = "proteinRaw",
 fun = MsCoreUtils::robustSummary)
```

```{r,warning=FALSE}
pe <- aggregateFeatures(pe,
 i = "peptidoform",
 fcol = "protein",
 na.rm = TRUE,
 name = "proteinRobust",
 fun = MsCoreUtils::robustSummary)
```

```{r}
#Only take pepforms that have a parent protein
#(when there is no global profiling dataset, this will be all peptidoforms)
pepWithProtein <- rowData(pe[["peptidoform"]])$protein %in% rownames(pe[["proteinRobust"]])
pepepWithProtein  <- pe[["peptidoform"]][pepWithProtein,]
pe <- addAssay(pe,pepepWithProtein,"pepformRel")
#normalisation for protein abundance step
assay(pe[["pepformRel"]]) <- assay(pe[["pepformRel"]]) - assay(pe[["proteinRobust"]])[rowData(pe[["pepformRel"]])$protein,colnames(assay(pe[["pepformRel"]]))]
boxplot(assay(pe[["pepformRel"]]))
```

### ptm summarisatie

Summarise peptidoforms to ptm level

#### Get ptm level intensity matrix

Get matrix for DPTMx -> get summarised intensity values

```{r}
ptms <- unique(as.vector(rowData(pe[["pepformRel"]])$ptm))
ptms <- ptms[-c(which(ptms==""))]
```


```{r, warning = F}
#For each ptm do
ptm_x_assay <- sapply(unique(ptms), function(i){ 
  #print(i)
  #Get current protein and mod from ptm
  #This is pretty straightforward here because there is only 1 ptm per peptidoform
  ptm_sub <- pe[["pepformRel"]][grepl(i, rowData(pe[["pepformRel"]])$ptm, fixed = T),]
  #ptm_sub <- filterFeatures(pe,~grepl(ptm,pattern=i, fixed=T))[["peptidoformNorm"]]
  #Get intensity values of those peptidoforms
  z <- assay(ptm_sub)
  z <- filter(as.data.frame(z), rowSums(is.na(z) | z==0) != ncol(z))
  #And summarise them to 1 row of intensity values: 1 value per sample for that ptm
  ptm_y <- try(MsCoreUtils::robustSummary(as.matrix(z)), silent = T)
  if (is(ptm_y, "try-error"))
  {ptm_y <- matrix(rep(NA,ncol(ptm_sub)), nrow = 1, 
                     dimnames = list(c(i), c(colnames(ptm_sub))))} 

  ptm_y
})
#Then we get the intensity assay on ptm level
ptm_x_assay <- t(ptm_x_assay)
#rownames(ptm_x_assay) <- ptms
```

#### Add to QFeatures object

Filter out ptms with all zero intensities or with the same intensity in every column

```{r}
print(paste(nrow(ptm_x_assay), "ptms before filtering"))
filtering <- (rowSums(ptm_x_assay != 0, na.rm=TRUE)) > 0 & 
  (rowSds(ptm_x_assay, na.rm=TRUE) > 1e-4) & 
          (rowSums(is.na(ptm_x_assay)) < (ncol(ptm_x_assay)-1))
ptm_x_assay <- ptm_x_assay[filtering,]
print(paste(nrow(ptm_x_assay), "ptms after filtering"))
```

```{r}
all(rownames(colData(pe)) == colnames(ptm_x_assay))
rowdata <- data.frame(ptm = rownames(ptm_x_assay))
rowdata$protein <- sapply(str_split(rowdata$ptm, pattern="_"),function(x) x[1])
ptm_y_assay <- SummarizedExperiment(assays=as.matrix(ptm_x_assay), rowData=rowdata, colData=colData(pe))
```

```{r}
pe <- addAssay(pe, ptm_y_assay, "ptmRel")
```

## Start from MSstats summarised data

PTM
```{r}
#Get to wide format for msqrob
ptm_df <- pivot_wider(summarized_ptm$PTM$ProteinLevelData, id_cols = c("Protein"),
                         names_from = "originalRUN", values_from = "LogIntensities")
ptm_df <- ptm_df %>% dplyr::rename(PROTEIN_PTM = Protein)
ptm_df$PROTEIN =  paste(sapply(ptm_df$PROTEIN_PTM, function (x) {str_split(x, "_")[[1]][1]}))
```

PROTEIN

```{r}
#Get to wide format for msqrob
protein_df <- summarized_ptm$PROTEIN$ProteinLevelData %>% arrange(originalRUN) %>%
              pivot_wider(id_cols = c("Protein"),
                         names_from = "originalRUN", values_from = "LogIntensities")
nrow(protein_df)
```

-> There are few proteins left here, because there are few proteins that have a non-modified peptide identified for them. 

### Add to QFeatures object

```{r}
#order dataframe by protein, for the normalisation step
ptm_df = ptm_df[order(ptm_df$PROTEIN),]
ptm <- readQFeatures(ptm_df,ecol= 2:17,name="ptmMSstats", fnames = "PROTEIN_PTM")
prot <- readQFeatures(protein_df,ecol= 2:17,name="protein",fnames="Protein")
#y_assay <- SummarizedExperiment(assays=as.matrix(y_new), rowData=rowdata, colData=colData(pe[["peptidoformRaw"]]))
pe <- addAssay(pe, ptm[["ptmMSstats"]], name = "ptmMSstats")
pe <- addAssay(pe, prot[["protein"]], name = "proteinMSstats")
```

The MSstats data is already preprocessed

```{r}
boxplot(assay(pe[["ptmMSstats"]]))
```



MSstats input data normalisation (PTM)

### Normalize PTM log intensity by protein log intensity

```{r}
#Only take ptms that have a parent protien
ptmWithProteinEvalue <- rowData(pe[["ptmMSstats"]])$PROTEIN %in% rownames(pe[["proteinMSstats"]])
ptmM <- pe[["ptmMSstats"]][ptmWithProteinEvalue,]
pe <- addAssay(pe,ptmM,"ptmMSstatsRel")
assay(pe[["ptmMSstatsRel"]]) <- assay(pe[["ptmMSstatsRel"]]) - assay(pe[["proteinMSstats"]])[rowData(pe[["ptmMSstatsRel"]])$PROTEIN,colnames(assay(pe[["ptmMSstatsRel"]]))]
boxplot(assay(pe[["ptmMSstatsRel"]]))
```

```{r}
pe
```

peptidoformRaw = raw peptidoform (log) intensities 
peptidoform = center median normalised peptidoform intensities
ptmMSstats = ptm data according to MSstats summarisation
proteinMSstats = protein data according to MSstats
proteinRaw = peptidoformRaw -> protein
proteinRobust = peptidoform -> protein
pepformRel = peptidoform normalised for protein abundance
ptmRel = ptm normalised for protein abundance when starting from raw data (typical msqrob2PTM workflow)
ptmMSstatsRel = ptm normalised for protein abundance when starting from MSstats summarisation

```{r}
nrow(assay(pe[["ptmMSstatsRel"]]))
nrow(assay(pe[["ptmRel"]]))
```

## Analysis

### Peptidoform analysis

Only possible when starting from the raw data

#### Hypothesistest for each contrast

```{r, warning=FALSE}
pe <- msqrob2::msqrob(object = pe, i = "pepformRel", formula = ~group + (1|biorep), robust = F, overwrite = T)
getCoef(rowData(pe[["pepformRel"]])$msqrobModels[[2]])
```


```{r}
contrasts <- c("groupCombo", "groupCCCP", "groupUSP30_OE", "groupCombo - groupCCCP",
               "groupCCCP - groupUSP30_OE", "groupCombo - groupUSP30_OE")
```

```{r}
L <- makeContrast(c("groupCombo = 0", 
                    "groupCCCP = 0", 
                    "groupUSP30_OE = 0", 
                    "groupCombo - groupCCCP = 0",
                    "groupCCCP - groupUSP30_OE = 0", 
                    "groupCombo - groupUSP30_OE = 0"), 
                  parameterNames = rowData(pe[["pepformRel"]])$msqrobModels[[2]] %>%
                  getCoef %>%
                  names)
pe <- hypothesisTest(object = pe, i = "pepformRel", contrast = L, overwrite=T)
```


#### Volcanoplot 

```{r}
volcano <- list()
for (contrast in contrasts){
volcano[[contrast]] <- rowData(pe[["pepformRel"]])[[contrast]]%>%
            ggplot(aes(x = logFC, y = -log10(pval), 
                   color = adjPval < 0.05,
                   annotation=rowData(pe[["pepformRel"]])[,3])) +
  geom_point(cex = 2.5) +
  scale_color_manual(values = alpha(c("black", "red"), 0.5)) + 
  theme_minimal() +
  ylab("-log10(pvalue)") +
  ggtitle(contrast)
}
volcano
```

#### Significant peptidoforms for each contrast

```{r}
tables <- list()
for (contrast in contrasts){
sigTable <- rowData(pe[["pepformRel"]])[[contrast]]
if(nrow(sigTable %>%
  na.exclude %>%
  filter(adjPval<0.05)) > 0){
sigTable <- sigTable %>%
  na.exclude %>%
  filter(adjPval<0.05) %>%
  arrange(pval) %>%
  mutate(
    se = format(se, digits = 2), 
    df = format(df, digits =2),
    t = format(t, digits = 2),
    adjPval = format(adjPval, digits = 3),
    rank = 1:length(logFC) 
  ) 
sigTable_print <- sigTable %>% mutate(peptidoform = rownames(sigTable)) %>% gt() %>% tab_header(title = md(contrast))
tables[[contrast]] <- sigTable
}
#knitr::kable(sigTable, caption = contrast)
}
knitr::kable(tables)
```

### PTM level analysis



```{r}
pe <- msqrob2::msqrob(object = pe, i = "ptmRel", formula = ~group+(1|biorep), robust=TRUE, overwrite = T)
pe <- msqrob2::msqrob(object = pe, i = "ptmMSstatsRel", formula = ~group+(1|biorep), robust=TRUE, overwrite = T)

pe <- hypothesisTest(object = pe, i = "ptmMSstatsRel", contrast = L, overwrite = T)
pe <- hypothesisTest(object = pe, i = "ptmRel", contrast = L, overwrite = T)
```

### Significant ptms for each contrast

```{r}
volcano <- list()
for (contrast in contrasts){
volcano[[contrast]] <- rowData(pe[["ptmRel"]])[[contrast]]%>%
            ggplot(aes(x = logFC, y = -log10(pval), 
                   color = adjPval < 0.05,
                   annotation=rowData(pe[["ptmRel"]])[,3])) +
  geom_point(cex = 2.5) +
  scale_color_manual(values = alpha(c("black", "red"), 0.5)) + 
  theme_minimal() +
  ylab("-log10(pvalue)") +
  ggtitle(contrast)
}
volcano
```

```{r}
tables <- list()
for (contrast in contrasts){
sigTable <- rowData(pe[["ptmRel"]])[[contrast]]
if(nrow(sigTable %>%
  na.exclude %>%
  filter(adjPval<0.05))>0){
sigTable <- sigTable %>%
  na.exclude %>%
  filter(adjPval<0.05) %>%
  arrange(pval) %>%
  mutate(
    se = format(se, digits = 2), 
    df = format(df, digits =2),
    t = format(t, digits = 2),
    adjPval = format(adjPval, digits = 3),
    rank = 1:length(logFC) 
  ) 
sigTable_print <- sigTable %>% mutate(peptidoform = rownames(sigTable)) %>% gt() %>% tab_header(title = md(contrast))
tables[[contrast]] <- sigTable
}
}
knitr::kable(tables)
```

```{r}
tables <- list()
for (contrast in contrasts){
sigTable <- rowData(pe[["ptmMSstatsRel"]])[[contrast]]
if(nrow(sigTable %>%
  na.exclude %>%
  filter(adjPval<0.05))>0){
sigTable <- sigTable %>%
  na.exclude %>%
  filter(adjPval<0.05) %>%
  arrange(pval) %>%
  mutate(
    se = format(se, digits = 2), 
    df = format(df, digits =2),
    t = format(t, digits = 2),
    adjPval = format(adjPval, digits = 3),
    rank = 1:length(logFC) 
  ) 
sigTable_print <- sigTable %>% mutate(peptidoform = rownames(sigTable)) %>% gt() %>% tab_header(title = md(contrast))
tables[[contrast]] <- sigTable
}
}
knitr::kable(tables)
```

```{r eval=FALSE, include=FALSE}
save(pe, file = "C:/Users/Nina/OneDrive - UGent/Documenten/Doctoraat/msqrobPTM/biological ubiquitin dataset/pe_ubiquitin_msqrob.rda")
```

Significant ptms from MSstats analysis

```{r}
for (contrast in model_ptm$ADJUSTED.Model$Label %>% unique){
  print(contrast)
  print(model_ptm$ADJUSTED.Model %>% filter(Label==contrast, adj.pvalue <= 0.05, Adjusted == T))
}
```

We can see a lot higher number of significant ptms here.