---
title: "vignette msqrobPTM no non-enriched dataframe"
output:
  html_document:
    code_download: yes
    theme: cosmo
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
    highlight: tango
    number_sections: yes
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: '4'
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/Nina/OneDrive - UGent/Documenten/Doctoraat/msqrobPTM paper/vignette/")
```

```{r message=FALSE}
library(QFeatures)
library(msqrob2)
library(tidyverse)
library(gt)
library(plotly)
library(stageR)
library(poolr)
library(RColorBrewer)
library(seqinr)
library(stringr)
library(ExploreModelMatrix)
```


# Import data

Here we have MaxQuant output, we will work with the evidence file. 
Any output of a search engine/quantification algorithm that contains the MS2 intensities of every peptidoform or psm per sample can be used here.

```{r}
df <- read.csv("evidence.txt",header=TRUE, sep = ",")
#filter out internal standard proteins
df <- df %>% filter(!grepl("Sigma", df$Proteins))
#filter out phospho sites with less than 75% probability
filter_rows <- sapply(df$Phospho..STY..Probabilities, function(x){
  if (!x==""){
    prob <- str_extract_all(x,  "(?<=\\().+?(?=\\))")[[1]]
    if (!any(as.double(prob) > 0.75)){return(which(df$Phospho..STY..Probabilities == x))}
  }
}, USE.NAMES = F)
filter_rows <- unique(unlist(filter_rows))
nrow(df)
df <- df[-filter_rows,]
nrow(df)
#transorm dataframe into wide format for input into QFeatures
#We need at least information about the pepform sequence, modifications, parent protein and intensity values
df_wide <- pivot_wider(df, id_cols = c("Sequence", "Modifications", "Modified.sequence", "Proteins", "Leading.proteins",
                                        "Reverse", "Potential.contaminant", "Protein.group.IDs", "Leading.razor.protein"),
                       names_from = "Raw.file", names_prefix = "Intensity_", values_from = "Intensity", values_fn = max)
```


```{r}
#get columns which contain the intensity values
ecols <- grep("Intensity",colnames(df_wide))
#order dataframe by protein, for the normalisation stept
df_wide = df_wide[order(df_wide$Leading.razor.protein),]
#read into QFeatures format while specifying intensity columns
pe <- readQFeatures(df_wide,ecol= ecols,name="peptidoformRaw", fnames = "Modified.sequence")
```

## Experimental Layout

We need to add information about the experimental design into the QFeatures object.
In this example, this information is present in an excel file.

```{r}
metadata <- read.csv("metadata.csv", sep = ";")
```

We then fill out the colData of the QFeatures object by extracting this information from the file we just read in.
This will later be used when doing the statistical modeling.

```{r}
colData(pe)$file <- sapply(str_split(rownames(colData(pe)), "_"), function(x) x[[2]])
rownames(metadata) <- metadata$Sample
colData(pe)$condition <- metadata[colData(pe)$file,]$Condition
colData(pe)$condition <- sapply(colData(pe)$condition, function(x) 
  ifelse(grepl("A", x), "A", "B"))
colData(pe)$subset <- metadata[colData(pe)$file,]$Subset
colData(pe)
```


# Preprocessing

### Handling overlapping protein groups (optional)
In our approach a peptide can map to multiple proteins, as long as there is none
of these proteins present in a smaller subgroup.

```{r}
nrow(assay(pe[["peptidoformRaw"]]))
pe[["peptidoformRaw"]] <- pe[["peptidoformRaw"]][rowData(pe[["peptidoformRaw"]])$Proteins %in%
                                           smallestUniqueGroups(rowData(pe[["peptidoformRaw"]])$Proteins), ]
nrow(assay(pe[["peptidoformRaw"]]))
```

### preprocessing

We then filter out decoys and contaminants and filter out any peptidoforms that have fewer than 2 associated intensity values.
We also perform a log transformation on the intensity values.

```{r}
rowData(pe[["peptidoformRaw"]])$nNonZero <- rowSums(assay(pe[["peptidoformRaw"]]) > 0, na.rm = T)
pe <- zeroIsNA(pe, i = "peptidoformRaw")
pe <- pe[rowData(pe[["peptidoformRaw"]])$nNonZero>2,,]
pe <- filterFeatures(pe, ~Reverse != "+")
pe <- filterFeatures(pe, ~Potential.contaminant != "+")
pe <- logTransform(pe, base = 2, i = "peptidoformRaw", name = "peptidoformLog")
pe <- normalize(pe, method = "center.median", i = "peptidoformLog", name = "peptidoform")
colData(pe[["peptidoform"]]) <- colData(pe)
```


## Normalisation via robust summarisation

We will now perform a normalisation of the peptidoform values.
For this, we need the summarised intensity values for every protein present in the dataset, 
which we obtain by using the aggregateFeatures function.
This function takes all peptidoforms from one protein and summarises its intensity values into one intensity value per sample for that protein.
Here, the function with which it does this is the robustSummary function from the MsCoreUtils package which utilises robust regression.

```{r,warning=FALSE}
pe <- aggregateFeatures(pe,
 i = "peptidoform",
 fcol = "Leading.razor.protein",
 na.rm = TRUE,
 name = "proteinRobust",
 fun = MsCoreUtils::robustSummary)
```

We then use these intensity values for each protein to normalise our dataframe.
For each peptidoform, from their corresponding intensity values, the summarised parent protein intensity value (per sample) is subtracted. 

```{r}
#Only take pepforms that have a parent protein
#(when there is no global profiling dataset, this will be all peptidoforms)
pepWithProtein <- rowData(pe[["peptidoform"]])$Proteins %in% rownames(pe[["proteinRobust"]])
pePepWithProtein  <- pe[["peptidoform"]][pepWithProtein,]
pe <- addAssay(pe,pePepWithProtein,"pepformRel")
#normalisation for protein abundance step
assay(pe[["pepformRel"]]) <- assay(pe[["pepformRel"]]) - assay(pe[["proteinRobust"]])[rowData(pe[["pepformRel"]])$Leading.razor.protein,colnames(assay(pe[["pepformRel"]]))]
boxplot(assay(pe[["pepformRel"]]))
```

# Differential peptidoform usage (DPFU)

We first check whether there are differentially used peptidoforms present.

## Hypothesistest for each contrast

We model the peptidoform level expression values using msqrob. By default msqrob2 estimates the model parameters using robust regression.

We will model the data with the groups incoded in the colData. We want to know whether there is difference in peptidoform usage between condition A and B and whether that difference is different for the different subsets. We can specify this model by using a formula with the factor 'condition' and 'subset' as its predictors: formula = ~condition*subset. Note that we include the interaction effect by using an asterisk in the formula. Also note that a formula always starts with a tilde ‘~’.

```{r}
#We specify at least the object, the assay and the formula in the function
pe <- msqrob2::msqrob(object = pe, i = "pepformRel", formula = ~condition*subset, robust=FALSE)
#We want to see the parameter names of the model
rowData(pe[["pepformRel"]])$msqrobModels[[3]] %>%
                  getCoef
```

conditionA is the reference class for condition and subset x is the reference class for subset So the mean log2 expression for samples from condition A and subset x is ‘(Intercept)'. The mean log2 expression for subset x samples from condition B is’(Intercept)+conditionB’. Hence, the average log2 fold change between condition A and condition B for samples from subset x is modeled using the parameter ‘conditionB’. Thus, we assess the contrast ‘conditionB=0’ with our statistical test.
The average log2 fold change between conditionA and conditionB for samples from subset y is modeled using the parameter ‘conditionB + conditionB:subsety’. Thus, we assess the contrast ‘conditionB + conditionB:subsety=0’ with our statistical test
The difference between abovementioned contrasts is modeled using the interaction parameter: conditionB:subsety
We can also assess an average contrast to see the average difference between condition A and B across subsets: conditionB + 0.5 * conditionB:subsety = 0

```{r}
contrasts <- c("conditionB", "conditionB + conditionB:subsety", "conditionB:subsety", "conditionB + 0.5 * conditionB:subsety")
L <- makeContrast(c("conditionB = 0",
                    "conditionB + conditionB:subsety = 0",
                    "conditionB:subsety = 0",
                    "conditionB + 0.5 * conditionB:subsety = 0"),
                  parameterNames = rowData(pe[["pepformRel"]])$msqrobModels[[3]] %>%
                  getCoef %>%
                  names)
#the default multiple testing correction is Benjamini Hochberg FDR correction
pe <- hypothesisTest(object = pe, i = "pepformRel", contrast = L)
```


### Volcanoplot 

```{r}
volcano <- list()
for (contrast in contrasts){
volcano[[contrast]] <- rowData(pe[["pepformRel"]])[[contrast]]%>%
            ggplot(aes(x = logFC, y = -log10(pval), 
                   color = adjPval < 0.05,
                   annotation=rowData(pe[["pepformRel"]])[,3])) +
  geom_point(cex = 2.5) +
  scale_color_manual(values = alpha(c("black", "red"), 0.5)) + 
  theme_minimal() +
  ylab("-log10(pvalue)") +
  ggtitle(contrast) +
  xlim(-5,5)
}
volcano
```


### Significant peptidoforms for each contrast


```{r}
tables <- list()
for (contrast in contrasts){
sigTable <- rowData(pe[["pepformRel"]])[[contrast]]
if(nrow(sigTable %>%
  na.exclude %>%
  filter(adjPval<0.05)) > 0){
sigTable <- sigTable %>%
  na.exclude %>%
  filter(adjPval<0.05) %>%
  arrange(pval) %>%
  mutate(
    se = format(se, digits = 2), 
    df = format(df, digits =2),
    t = format(t, digits = 2),
    adjPval = format(adjPval, digits = 3),
    rank = 1:length(logFC) 
  ) 
sigTable_print <- sigTable %>% mutate(peptidoform = rownames(sigTable)) %>% gt() %>% tab_header(title = md(contrast))
tables[[contrast]] <- sigTable

}
}
knitr::kable(tables)
```

We here find all significant peptidoforms per contrast in a table.
The peptidoforms are sorted based on the adjusted pvalues

# DPTM

## ptm summarisation

In this step we will summarise the peptidoform level data to ptm level data.

### Get location of modification in protein (optional)

For this we need the location information of each modification in the protein.
When this information is not provided, but we do have the fasta file, we can add this information ourselves.

Use this function carefully when not using MaxQuant data.
Ideally, the location should already have been provided in the input file. 
In that case, this step can be skipped.

```{r}
fasta <- "human.fasta"
parsed_fasta <- read.fasta(file = fasta, seqtype = "AA", as.string = T)
```

  
```{r}
#modified sequence column contains _ that hinders the location finding
rowData(pe[["pepformRel"]])$modified_sequence <- gsub("_", "", rowData(pe[["pepformRel"]])$Modified.sequence)
```


```{r}
get_ptm_location <- function(feature, data, fasta, mod_column = "Modifications", 
                             peptide_seq_column = "Sequence", mod_seq_column = "modified_sequence", 
                             protein_column = "Leading.razor.protein", split = ",", collapse = ", "){

  prot <- data[feature,][[protein_column]]
  pep_seq <- data[feature,][[peptide_seq_column]]
  mod_seq <- data[feature,][[mod_seq_column]]
  prot_seq <- fasta[[prot]][1]

  #find location of peptide in protein
  #add -1 here, so that the addition of the location later on is correct
  pep_location <- unlist(lapply(gregexpr(pattern = pep_seq, prot_seq), min)) - 1
  final_mod <- c()
  j <- mod_seq
  #go over the modifications inside the modified sequence
  for(mod in regmatches(mod_seq, gregexpr("\\(.*?\\)\\)", mod_seq, perl=T))[[1]]){
    #find location of modification in peptide
    mod_location <- unlist(lapply(gregexpr(pattern = mod, j, fixed = T), min))
    #get location in protein (-1 because else it gives you the location after because of the presence of the modification in the string)
    location <- mod_location + pep_location -1
    #add location to modification
    mod_ <- paste(mod, location)
    #save modification
    final_mod <- c(final_mod, mod_)
    #now remove current modification from the sequence, so that we can continue to the next mod
    str_sub(j, mod_location, nchar(mod)+mod_location-1) <- ""
  }
  return(paste(final_mod, collapse = collapse))
}
```

```{r}
rowData(pe[["pepformRel"]])$mod <- sapply(rownames(rowData(pe[["pepformRel"]])), function(x){
    get_ptm_location(x, rowData(pe[["pepformRel"]]), parsed_fasta)
}, USE.NAMES = F)
```


### Get ptm level intensity matrix

In our case, a ptm is a unique protein - modification + location combination.

```{r}
#Add ptm variable = protein + modification 
#Do this only for peptidoforms with modifications
rowData(pe[["pepformRel"]])$ptm <- ifelse(rowData(pe[["pepformRel"]])$mod != "",
                                          #if there is mod, add ptm
                                               mapply(function(x, y) {
vector2 <- strsplit(y, ", ")[[1]]
result <- lapply(vector2, function(i) paste(x, i, sep = "-"))
final_result <- paste(unlist(result), collapse = ", ")
}, rowData(pe[["pepformRel"]])$Leading.razor.protein, rowData(pe[["pepformRel"]])$mod), 
#else add empty string
                                               "")
```

First, we need a vector of all of these unique ptms.

```{r}
#Get all unique PTMs
ptms <- unique(unlist(strsplit(rowData(pe[["pepformRel"]])$ptm, split = ", ", fixed = TRUE)))
```

For every unique ptm, we take all its associated peptidoforms and aggregate all its intensity values per sample into one intensity value per sample for that ptm.

```{r}
#For each ptm do
ptm_x_assay <- sapply(seq(1:length(ptms)), function(i){ 
  x <- ptms[i]
  #Get current protein and mod from ptm
  prot <- str_split(x, "-", 2)[[1]][1]
  current_ptm <- str_split(x, "-", 2)[[1]][2]
  #filter on that protein and on that mod to obtain all peptidoforms that correspond to the ptm
  pe_sub <- pe[["pepformRel"]][grepl(prot, rowData(pe[["pepformRel"]])$Leading.razor.protein, fixed = T),]
  ptm_sub <- pe_sub[grepl(current_ptm, rowData(pe_sub)$mod, fixed = T),]
  #Get intensity values of those peptidoforms
  y <- assay(ptm_sub)
  #And summarise them to 1 row of intensity values: 1 value per sample for that ptm
  ptm_y <- try(MsCoreUtils::robustSummary(y), silent = T)
  if (is(ptm_y, "try-error")){
    ptm_y <- rep(NA, ncol(y))}
  ptm_y
})
#Then we get the intensity assay on ptm level
ptm_x_assay <- t(ptm_x_assay)
rownames(ptm_x_assay) <- ptms
```

### Make new QFeatures object

Filter out ptms with all zero intensities or with the same intensity in every column
Then add the newly made assay to the existing pe, filling out its rowData and colData

```{r}
print(paste(nrow(ptm_x_assay), "ptms before filtering"))
filtering <- (rowSums(ptm_x_assay != 0, na.rm=TRUE)) > 0 & 
  (rowSds(ptm_x_assay, na.rm=TRUE) > 1e-4) & 
          (rowSums(is.na(ptm_x_assay)) < (ncol(ptm_x_assay)-1))
ptm_x_assay <- ptm_x_assay[filtering,]
print(paste(nrow(ptm_x_assay), "ptms after filtering"))
```

```{r}
all(rownames(colData(pe)) == colnames(ptm_x_assay))
rowdata <- data.frame(ptm = rownames(ptm_x_assay))
rowdata$Protein_Family <- sapply(str_split(rowdata$ptm, pattern="_"),function(x) x[1])
ptm_y_assay <- SummarizedExperiment(assays=as.matrix(ptm_x_assay), rowData=rowdata, colData=colData(pe))
```


```{r}
pe <- addAssay(pe, ptm_y_assay, "ptmRel")
```


## Hypothesistest for each contrast

We again model our data and assess the same contrasts as for the peptidoform analysis.

```{r}
pe <- msqrob2::msqrob(object = pe, i = "ptmRel", formula = ~condition*subset, robust=TRUE, overwrite = T)
rowData(pe[["ptmRel"]])$msqrobModels[[3]] %>%
                  getCoef
```


```{r}
contrasts <- c("conditionB", "conditionB + conditionB:subsety", "conditionB:subsety", "conditionB + 0.5 * conditionB:subsety")
L <- makeContrast(c("conditionB = 0",
                    "conditionB + conditionB:subsety = 0",
                    "conditionB:subsety = 0",
                    "conditionB + 0.5 * conditionB:subsety = 0"),
                  parameterNames = rowData(pe[["ptmRel"]])$msqrobModels[[3]] %>%
                  getCoef %>%
                  names)
pe <- hypothesisTest(object = pe, i = "ptmRel", contrast = L, overwrite = T)
```



### Volcanoplot 


```{r}
volcanos <- list()
for (contrast in contrasts){
library(plotly)
volcanos[[contrast]] <- 
  rowData(pe[["ptmRel"]])[[contrast]]%>%
  ggplot(aes(x = logFC, y = -log10(pval), 
             color = adjPval < 0.05,
             annotation=rowData(pe[["ptmRel"]])[,3])) +
  geom_point(cex = 2.5) +
  scale_color_manual(values = alpha(c("black", "red"), 0.5)) + theme_minimal() +
  ylab("-log10(pvalue)") +
  ggtitle(contrast)
}
volcanos
```


### Significant ptms for each contrast


```{r}
tables <- list()
for (contrast in contrasts){
sigTable <- rowData(pe[["ptmRel"]])[[contrast]]
if(nrow(sigTable %>%
  na.exclude %>%
  filter(adjPval<0.05))>0){
sigTable <- sigTable %>%
  na.exclude %>%
  filter(adjPval<0.05) %>%
  arrange(pval) %>%
  mutate(
    se = format(se, digits = 2), 
    df = format(df, digits =2),
    t = format(t, digits = 2),
    adjPval = format(adjPval, digits = 3),
    rank = 1:length(logFC) 
  ) 
sigTable_print <- sigTable %>% mutate(peptidoform = rownames(sigTable)) %>% gt() %>% tab_header(title = md(contrast))
tables[[contrast]] <- sigTable
}
}
knitr::kable(tables)
```

We here find all significant PTMs per contrast in a table.
The ptms are sorted based on the adjusted pvalues
Because we only took a subset of a larger dataset for speed of calculation in this example, very few PTMs are found significant.