---
title: "phosphoCSF GP mock new new workflow"
output: html_document
date: "2023-01-09"
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/Nina/OneDrive - UGent/Documenten/Doctoraat/msqrobPTM paper/mock analyses/")
```

```{r}
library(dplyr)
library(tidyr)
library(tibble)
library(msqrob2)
library(QFeatures)
library(stringr)
library(seqinr)
```

```{r}
metadata <- read.csv("Experimental Design.csv")
```

```{r}
df <- read.csv("evidence.txt",header=TRUE, sep = ",")
#Only take B and y samples
samples <- metadata %>% filter(Subset=="y",
                               grepl("B", Condition)) %>%
           pull(File) %>% gsub(".raw", "", .)
#filter out QC stuff
df <- df %>% filter(grepl(paste(samples, collapse = "|"), df$Raw.file),
                    !grepl("Sigma", df$Proteins))
#Now get format into wide format
df_wide <- pivot_wider(df, id_cols = c("Sequence", "Modifications", "Modified.sequence", "Proteins", "Leading.proteins",
                                        "Reverse", "Potential.contaminant", "Protein.group.IDs", "Leading.razor.protein"),
                       names_from = "Raw.file", names_prefix = "Intensity_", values_from = "Intensity", values_fn = max)
```

```{r}
ecols <- grep("Intensity",colnames(df_wide))
#order dataframe by protein, for the normalisation step
df_wide = df_wide[order(df_wide$Leading.proteins),]
pe <- readQFeatures(df_wide,ecol= ecols,name="peptidoformRaw", fnames = "Modified.sequence")
```


## Experimental Layout

```{r}
colData(pe)$file <- sapply(str_split(rownames(colData(pe)), "_"), function(x) x[[2]])
rownames(metadata) <- metadata$File
colData(pe)$condition <- metadata[colData(pe)$file,]$Condition
colData(pe)$condition <- sapply(colData(pe)$condition, function(x) 
  ifelse(grepl("A", x), "A", "B"))
colData(pe)$subset <- metadata[colData(pe)$file,]$Subset
colData(pe)
```


## Protein dataset

```{r}
df_prot <- read.csv("non_enriched_samples/evidence.txt",header=TRUE, sep = ",")
rawfilenames <- read.csv("rawfilenames_change.csv")
#I can ignore the pool samples (those were just used for QC purposes), same for Sigma proteins
samples_prot <- rawfilenames[rawfilenames$ptm.dataset %in% samples,]$protein.dataset
df_prot <- df_prot %>% filter( !grepl("Sigma", df_prot$Proteins),
                    #corresponding protein files for B y
                    grepl(paste(samples_prot, collapse = "|"), df_prot$Raw.file))

#Now get format into wide format
df_prot_wide <- pivot_wider(df_prot, id_cols = c("Sequence", "Modifications", "Modified.sequence", "Proteins", "Leading.proteins",
                                        "Reverse", "Potential.contaminant", "Protein.group.IDs", "Leading.razor.protein"),
                       names_from = "Raw.file", names_prefix = "Intensity_", values_from = "Intensity", values_fn = max)
```

```{r}
ecols <- grep("Intensity",colnames(df_prot_wide))
#order dataframe by protein, for the normalisation step
df_prot_wide = df_prot_wide[order(df_prot_wide$Leading.razor.protein),]
prot <- readQFeatures(df_prot_wide,ecol= ecols,name="proteinRaw", fnames = "Modified.sequence")
```

Total protein overlap

```{r}
prot_GP <- unique(df_prot_wide$Leading.razor.protein)
prot_P <- unique(df_wide$Leading.razor.protein)
#how many of the protein in enriched dataset are not in non-enriched dataset?
n_diff <- length(setdiff(prot_P, prot_GP))
n_diff/length(prot_P)
```

25% of proteins in the enriched dataset do not find a counterpart in the non-enriched dataset

### Experimental Layout

```{r}
metadata <- read.csv("non_enriched_samples/metadata.csv")
```

```{r}
colData(prot)$file <- sapply(str_split(rownames(colData(prot)), "_"), function(x) x[[2]])
metadata$File <- sapply(strsplit(metadata$File, ".", fixed = T), function(x) x[[1]])
metadata <- metadata %>% filter(!grepl(paste(setdiff(metadata$Condition, colData(pe)$conditionfull), collapse="|"), metadata$Condition))
colData(prot)$condition <- sapply(colData(prot)$file, function(x){
  metadata[metadata$File==x,]$Condition
}, USE.NAMES = F)
colData(prot)$subset <- sapply(colData(prot)$file, function(x){
  metadata[metadata$File==x,]$Subset
}, USE.NAMES = F)
colData(prot)$condition <- case_when(grepl("A", colData(prot)$condition) ~ "A",
                                        TRUE ~ "B")
```


```{r eval=FALSE, include=FALSE}
MSnbase::plotNA(assay(prot[["proteinRaw"]])) +
  xlab("Peptide index (ordered by data completeness)")
```


```{r}
nrow(assay(pe[["peptidoformRaw"]]))
pe[["peptidoformRaw"]] <- pe[["peptidoformRaw"]][rowData(pe[["peptidoformRaw"]])$Proteins %in%
                                  smallestUniqueGroups(rowData(pe[["peptidoformRaw"]])$Proteins), ]
nrow(assay(pe[["peptidoformRaw"]]))
```

## Preprocessing

```{r}
rowData(pe[["peptidoformRaw"]])$nNonZero <- rowSums(assay(pe[["peptidoformRaw"]]) > 0, na.rm = T)
pe <- pe[rowData(pe[["peptidoformRaw"]])$nNonZero>2,,]
pe <- zeroIsNA(pe, i = "peptidoformRaw")
pe <- filterFeatures(pe, ~Reverse != "+")
pe <- filterFeatures(pe, ~Potential.contaminant != "+")
pe <- logTransform(pe, base = 2, i = "peptidoformRaw", name = "peptidoformLog")
pe <- normalize(pe, i = "peptidoformLog", method = "center.median", name = "peptidoform")
colData(pe[["peptidoform"]]) <- colData(pe)
```

```{r}
rowData(prot[["proteinRaw"]])$nNonZero <- rowSums(assay(prot[["proteinRaw"]]) > 0, na.rm = T)
prot <- prot[rowData(prot[["proteinRaw"]])$nNonZero>2,,]
prot <- zeroIsNA(prot, i = "proteinRaw")
prot <- filterFeatures(prot, ~Reverse != "+")
prot <- filterFeatures(prot, ~Potential.contaminant != "+")
prot <- logTransform(prot, base = 2, i = "proteinRaw", name = "proteinLog")
prot <- normalize(prot, i = "proteinLog", method = "center.median", name = "protein")
colData(prot[["protein"]]) <- colData(prot)
```

## Aggregation to protein level

```{r,warning=FALSE}
prot <- aggregateFeatures(prot,
 i = "protein",
 fcol = "Leading.razor.protein",
 na.rm = TRUE,
 name = "proteinRobust",
 fun = MsCoreUtils::robustSummary)
```

## Normalize PTM log intensity by protein log intensity 

Protein dataset and ptm dataset need to have to same sample names to be able to couple them for the normalisation
 
```{r}
rawfilenames <- read.csv("rawfilenames_change.csv")
#get current protein colnames without "Intensity_"
current_colnames <- sapply(strsplit(colnames(assay(prot[["proteinRobust"]])), "_"), function(x) x[[2]])
#replace with corresponding ptm colnames
colnames(assay(prot[["proteinRobust"]], withDimnames = F)) <- 
  sapply(current_colnames, function(x){
                  ptm_colname <- rawfilenames[rawfilenames[["protein.dataset"]]==x,]$`ptm.dataset`
                  paste0("Intensity_", ptm_colname)
                }, USE.NAMES = F)
```

Do actual normalisation

```{r}
#Only take pepforms that have a parent protein
pepWithProtein <- rowData(pe[["peptidoform"]])$Leading.razor.protein %in% rownames(prot[["proteinRobust"]])
pepepWithProtein  <- pe[["peptidoform"]][pepWithProtein,]
pe <- addAssay(pe,pepepWithProtein,"pepformRel")
#normalisation for protein abundance step
assay(pe[["pepformRel"]]) <- assay(pe[["pepformRel"]]) - assay(prot[["proteinRobust"]], withDimnames = F)[rowData(pe[["pepformRel"]])$Leading.razor.protein,colnames(assay(pe[["pepformRel"]]))]
boxplot(assay(pe[["pepformRel"]]))
```


## PTM level

```{r}
fasta <- "human.fasta"
parsed_fasta <- read.fasta(file = fasta, seqtype = "AA", as.string = T)
#modified sequence kolom contain _ that do not matter, but prevent determining location 
rowData(pe[["pepformRel"]])$modified_sequence <- gsub("_", "", rowData(pe[["pepformRel"]])$Modified.sequence)
```

We need to know the location of the modification in the protein  

```{r}
get_ptm_location <- function(feature, data, fasta, mod_column = "Modifications", 
                             peptide_seq_column = "Sequence", mod_seq_column = "modified_sequence", 
                             protein_column = "Leading.razor.protein", split = ",", collapse = ", "){

  prot <- data[feature,][[protein_column]]
  pep_seq <- data[feature,][[peptide_seq_column]]
  mod_seq <- data[feature,][[mod_seq_column]]
  prot_seq <- fasta[[prot]][1]

  #find location of peptide in protein
  #add -1 here, so that the addition of the location later on is correct
  pep_location <- unlist(lapply(gregexpr(pattern = pep_seq, prot_seq), min)) - 1
  final_mod <- c()
  j <- mod_seq
  #go over the modifications inside the modified sequence
  for(mod in regmatches(mod_seq, gregexpr("\\(.*?\\)\\)", mod_seq, perl=T))[[1]]){
    #find location of modification in peptide
    mod_location <- unlist(lapply(gregexpr(pattern = mod, j, fixed = T), min))
    #get location in protein (-1 because else it gives you the location after because of the presence of the modification in the string)
    location <- mod_location + pep_location -1
    #add location to modification
    mod_ <- paste(mod, location)
    #save modification
    final_mod <- c(final_mod, mod_)
    #now remove current modification from the sequence, so that we can continue to the next mod
    str_sub(j, mod_location, nchar(mod)+mod_location-1) <- ""
  }
  return(paste(final_mod, collapse = collapse))
}
```

```{r}
rowData(pe[["pepformRel"]])$mod <- sapply(rownames(rowData(pe[["pepformRel"]])), function(x){
    get_ptm_location(x, rowData(pe[["pepformRel"]]), parsed_fasta)
}, USE.NAMES = F)
```


```{r}
#Add ptm variable = protein + modification
rowData(pe[["pepformRel"]])$ptm <- ifelse(rowData(pe[["pepformRel"]])$mod != "",
                                               paste(rowData(pe[["pepformRel"]])$Leading.razor.protein, rowData(pe[["pepformRel"]])$mod, sep="_"), "")
```


Get all unique ptms present in the dataset

```{r}
prots <- unique(rowData(pe[["pepformRel"]])$Leading.razor.protein)
#Do for each protein
ptms <- sapply(prots, function(i) {
  pe_sub <- pe[["pepformRel"]][rowData(pe[["pepformRel"]])$Leading.razor.protein %in% i,]
  #pe_sub <- filterFeatures(pe,~grepl(Leading.razor.protein,pattern=i,fixed = T))
  #Get all unique modification present on that protein
  mods <- unique(unlist(strsplit(rowData(pe_sub)$mod, split = ", ", fixed = TRUE)))
  #Add protein info to mods
  ptm <- paste(rep(i, length(mods)), mods)
  #return all the protein-mods combinations
  ptm
})
ptms <- as.vector(unlist(ptms))
```

Get ptm level intensity matrix

```{r}
pe <- zeroIsNA(pe, "pepformRel")

#For each ptm do
ptm_x_assay <- sapply(seq(1:length(ptms)), function(i){ 
  x <- ptms[i]
  #Get current protein and mod from ptm
  prot <- str_split(x, " ", 2)[[1]][1]
  current_ptm <- str_split(x, " ", 2)[[1]][2]
  #filter on that protein and on that mod to obtain all peptidoforms that correspond to the ptm
  pe_sub <- pe[["pepformRel"]][rowData(pe[["pepformRel"]])$Leading.razor.protein %in% prot,]
  ptm_sub <- pe_sub[grepl(current_ptm, rowData(pe_sub)$mod, fixed = T),]
  # pe_sub <- filterFeatures(pe,~grepl(Leading.razor.protein,pattern=prot, fixed = T))
  # ptm_sub <- filterFeatures(pe_sub,~grepl(mod,pattern=current_ptm, fixed=T))[["pepformRel"]]
  #Get intensity values of those peptidoforms
  y <- assay(ptm_sub)
  #And summarise them to 1 row of intensity values: 1 value per sample for that ptm
  ptm_y <- try(MsCoreUtils::robustSummary(as.matrix(y)), silent = T)
  if (is(ptm_y, "try-error"))
  {ptm_y <- matrix(rep(NA,ncol(ptm_sub)), nrow = 1, 
                     dimnames = list(c(i), c(colnames(ptm_sub))))} 

  ptm_y
})
#Then we get the intensity assay on ptm level
ptm_x_assay <- t(ptm_x_assay)
rownames(ptm_x_assay) <- ptms
```

Get ptm level intensity matrix into new QFeatures object

```{r}
ptm_y_assay <- as.matrix.data.frame(ptm_x_assay)
rownames(ptm_y_assay) <- rownames(ptm_x_assay)
colnames(ptm_y_assay) <- colnames(assay(pe))
ptm <- readQFeatures(as.data.frame(ptm_y_assay),ecol= 1:ncol(ptm_y_assay),name="ptm")
ptm <- renamePrimary(ptm, rownames(colData(pe)))
colData(ptm) <- colData(pe)
rownames(ptm[["ptm"]]) <- rownames(ptm_x_assay)
rowData(ptm[["ptm"]])$Protein <- sapply(str_split(rownames(ptm[["ptm"]]), pattern=" ", n=2),function(x) x[1])
rowData(ptm[["ptm"]])$modification <- sapply(str_split(rownames(ptm[["ptm"]]), pattern=" ", n=2),function(x) x[2])
colData(ptm)$condition <- relevel(as.factor(colData(ptm)$condition), ref = "B")

print(nrow(assay(ptm[["ptm"]])))
ptm <- zeroIsNA(ptm, "ptm")
rowData(ptm[["ptm"]])$nNA <- nNA(ptm, 1)$nNArows$nNA
#ptm <- filterFeatures(ptm, ~nNA < 1)
print(nrow(assay(ptm[["ptm"]])))
```

# Simulations

### Peptidoform level

```{r warning=FALSE}
set.seed(645)
pvals <- matrix(NA,nrow=nrow(assay(pe[["pepformRel"]])),ncol=5)
adj_pvals <- matrix(NA,nrow=nrow(assay(pe[["pepformRel"]])),ncol=5)
fprs <- c()
L <- makeContrast(c("conditionA = 0"), parameterNames = c("(Intercept)", "conditionA"))

for (i in 1:5)
{
# if(i%%50==0){
#   print(i)
# }
#There are 26 samples
nSamp <- colData(pe) %>% rownames %>% unique() %>% length()
#We randomly choose 13 out of 26 samples
mocks <- sample(colData(pe) %>% rownames %>% unique(),nSamp%/%2)
colData(pe)$condition <- "B"
#those 9 get "A" assigned, rest gets B
colData(pe)[mocks,]$condition <- "A"
colData(pe)$condition <- relevel(as.factor(colData(pe)$condition), ref = "B")

colData(pe[["pepformRel"]]) <- colData(pe)

pe <- msqrob2::msqrob(object = pe, i = "pepformRel", formula = ~condition, robust=FALSE, overwrite = T)
pe <- hypothesisTest(object = pe, i = "pepformRel", contrast = L, overwrite = T)
pvals[,i] <- rowData(pe)[["pepformRel"]]$conditionA$pval
adj_pvals[,i] <- rowData(pe)[["pepformRel"]]$conditionA$adjPval
hist(pvals, breaks = 50, main = "Histogram of pvals - peptidoform level - no robust regression", cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5)
}

```



```{r warning=FALSE}
set.seed(745)
pvals <- matrix(NA,nrow=nrow(assay(pe[["pepformRel"]])),ncol=5)
adj_pvals <- matrix(NA,nrow=nrow(assay(pe[["pepformRel"]])),ncol=5)
L <- makeContrast(c("conditionA = 0"), parameterNames = c("(Intercept)", "conditionA"))

for (i in 1:5)
{
# if(i%%50==0){
#   print(i)
# }
#There are 18 samples
nSamp <- colData(pe) %>% rownames %>% unique() %>% length()
#We randomly choose 9 out of 18 samples
mocks <- sample(colData(pe) %>% rownames %>% unique(),nSamp%/%2)
colData(pe)$condition <- "B"
#those 9 get "A" assigned, rest gets B
colData(pe)[mocks,]$condition <- "A"
colData(pe)$condition <- relevel(as.factor(colData(pe)$condition), ref = "B")

colData(pe[["pepformRel"]]) <- colData(pe)

pe <- msqrob2::msqrob(object = pe, i = "pepformRel", formula = ~condition, robust=TRUE, overwrite = T)
pe <- hypothesisTest(object = pe, i = "pepformRel", contrast = L, overwrite = T)
pvals[,i] <- rowData(pe)[["pepformRel"]]$conditionA$pval
adj_pvals[,i] <- rowData(pe)[["pepformRel"]]$conditionA$adjPval
hist(pvals, breaks = 50, main = "Histogram of pvals - peptidoform level - robust regression")
}

```

### PTM level

```{r warning=FALSE}
set.seed(145)
pvals <- matrix(NA,nrow=nrow(assay(ptm[["ptm"]])),ncol=5)
adj_pvals <- matrix(NA,nrow=nrow(assay(ptm[["ptm"]])),ncol=5)
L <- makeContrast(c("conditionA = 0"), parameterNames = c("(Intercept)", "conditionA"))

for (i in 1:5)
{
# if(i%%50==0){
#   print(i)
# }
#There are 18 samples
nSamp <- colData(ptm) %>% rownames %>% unique() %>% length()
#We randomly choose 9 out of 18 samples
mocks <- sample(colData(ptm) %>% rownames %>% unique(),nSamp%/%2)
colData(ptm)$condition <- "B"
#those 9 get "A" assigned, rest gets B
colData(ptm)[mocks,]$condition <- "A"
colData(ptm)$condition <- relevel(as.factor(colData(ptm)$condition), ref = "B")

#colData(ptm[["ptm"]]) <- colData(ptm)

ptm <- msqrob2::msqrob(object = ptm, i = "ptm", formula = ~condition, robust=FALSE, overwrite = T)
ptm <- hypothesisTest(object = ptm, i = "ptm", contrast = L, overwrite = T)
pvals[,i] <- rowData(ptm)[["ptm"]]$conditionA$pval
adj_pvals[,i] <- rowData(ptm)[["ptm"]]$conditionA$adjPval
hist(pvals, breaks = 50, main = "Histogram of pvals - no robust regression")
}

```



```{r warning=FALSE}
set.seed(245)
pvals <- matrix(NA,nrow=nrow(assay(ptm[["ptm"]])),ncol=5)
adj_pvals <- matrix(NA,nrow=nrow(assay(ptm[["ptm"]])),ncol=5)
L <- makeContrast(c("conditionA = 0"), parameterNames = c("(Intercept)", "conditionA"))

for (i in 1:5)
{
# if(i%%50==0){
#   print(i)
# }
#There are 18 samples
nSamp <- colData(ptm) %>% rownames %>% unique() %>% length()
#We randomly choose 9 out of 18 samples
mocks <- sample(colData(ptm) %>% rownames %>% unique(),nSamp%/%2)
colData(ptm)$condition <- "B"
#those 9 get "A" assigned, rest gets B
colData(ptm)[mocks,]$condition <- "A"
colData(ptm)$condition <- relevel(as.factor(colData(ptm)$condition), ref = "B")
#colData(ptm[["ptm"]]) <- colData(ptm)

ptm <- msqrob2::msqrob(object = ptm, i = "ptm", formula = ~condition, robust=TRUE, overwrite = T)
ptm <- hypothesisTest(object = ptm, i = "ptm", contrast = L, overwrite = T)
pvals[,i] <- rowData(ptm)[["ptm"]]$conditionA$pval
adj_pvals[,i] <- rowData(ptm)[["ptm"]]$conditionA$adjPval
hist(pvals, breaks = 50, main = "Histogram of pvals - robust regression")
}
```


