---
title: "phospho mock no GP"
output:
  html_document:
    code_download: yes
    theme: cosmo
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
    highlight: tango
    number_sections: yes
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: '4'
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/Nina/OneDrive - UGent/Documenten/Doctoraat/msqrobPTM paper/mock analyses/")
```

```{r message=FALSE}
library(QFeatures)
library(msqrob2)
library(dplyr)
library(tidyr)
library(tibble)
library(gt)
library(plotly)
library(stageR)
library(poolr)
library(RColorBrewer)
library(seqinr)
library(stringr)
library(ExploreModelMatrix)
library(data.table)
```



# Import data

```{r}
metadata <- read.csv("Experimental Design.csv")
```

```{r}
df <- read.csv("evidence.txt",header=TRUE, sep = ",")
#I can ignore the sigma proteins (those were just used for QC purposes)
#Only take B and y samples
samples <- metadata %>% filter(Subset=="y",
                               grepl("B", Condition)) %>%
           pull(File) %>% gsub(".raw", "", .)
#filter out QC stuff
df <- df %>% filter(grepl(paste(samples, collapse = "|"), df$Raw.file),
                    !grepl("Sigma", df$Proteins))
#filter out phospho sites with less than 75% probability
#(other filtering options are possible)
filter_rows <- sapply(df$Phospho..STY..Probabilities, function(x){
  if (!x==""){
    prob <- str_extract_all(x,  "(?<=\\().+?(?=\\))")[[1]]
    if (!any(as.double(prob) > 0.75)){return(which(df$Phospho..STY..Probabilities == x))}
  }
}, USE.NAMES = F)
filter_rows <- unique(unlist(filter_rows))
nrow(df)
df <- df[-filter_rows,]
nrow(df)
#Now get format into wide format
df_wide <- pivot_wider(df, id_cols = c("Sequence", "Modifications", "Modified.sequence", "Proteins", "Leading.proteins",
                                        "Reverse", "Potential.contaminant", "Protein.group.IDs", "Leading.razor.protein"),
                                                                          #Is dat hier de beste manier om met 
                                                                          #gedupliceerde features om te gaan?
                       names_from = "Raw.file", names_prefix = "Intensity_", values_from = "Intensity", values_fn = max)
```


```{r}
ecols <- grep("Intensity",colnames(df_wide))
#order dataframe by protein, for the normalisation step
df_wide = df_wide[order(df_wide$Leading.razor.protein),]
pe <- readQFeatures(df_wide,ecol= ecols,name="peptidoformRaw", fnames = "Modified.sequence")
```


## Experimental Layout

```{r}
metadata <- read.csv("Experimental Design.csv")
```

```{r}
colData(pe)$file <- sapply(str_split(rownames(colData(pe)), "_"), function(x) x[[2]])
rownames(metadata) <- metadata$File
colData(pe)$condition <- metadata[colData(pe)$file,]$Condition
colData(pe)$condition <- sapply(colData(pe)$condition, function(x) 
  ifelse(grepl("A", x), "A", "B"))
colData(pe)$subset <- metadata[colData(pe)$file,]$Subset
colData(pe)
```


```{r}
MSnbase::plotNA(assay(pe[["peptidoformRaw"]])) +
  xlab("Peptide index (ordered by data completeness)")
```


# Preprocessing
```{r}
nrow(assay(pe[["peptidoformRaw"]]))
pe[["peptidoformRaw"]] <- pe[["peptidoformRaw"]][rowData(pe[["peptidoformRaw"]])$Proteins %in%
                                                   smallestUniqueGroups(rowData(pe[["peptidoformRaw"]])$Proteins), ]
nrow(assay(pe[["peptidoformRaw"]]))
```


```{r}
rowData(pe[["peptidoformRaw"]])$nNonZero <- rowSums(assay(pe[["peptidoformRaw"]]) > 0, na.rm = T)
pe <- pe[rowData(pe[["peptidoformRaw"]])$nNonZero>2,,]
pe <- zeroIsNA(pe, i = "peptidoformRaw")
pe <- filterFeatures(pe, ~Reverse != "+")
pe <- filterFeatures(pe, ~Potential.contaminant != "+")
pe <- logTransform(pe, base = 2, i = "peptidoformRaw", name = "peptidoformLog")
pe <- QFeatures::normalize(pe, method = "center.median", i = "peptidoformLog", name = "peptidoform")
colData(pe[["peptidoform"]]) <- colData(pe)
```


## Normalisation via robust summarisation


```{r,warning=FALSE}
pe <- aggregateFeatures(pe,
 i = "peptidoform",
 fcol = "Leading.razor.protein",
 na.rm = TRUE,
 name = "proteinRobust",
 fun = MsCoreUtils::robustSummary)
```

```{r}
#Only take pepforms that have a parent protein
#(when there is no global profiling dataset, this will be all peptidoforms)
pepWithProtein <- rowData(pe[["peptidoform"]])$Proteins %in% rownames(pe[["proteinRobust"]])
pePepWithProtein  <- pe[["peptidoform"]][pepWithProtein,]
pe <- addAssay(pe,pePepWithProtein,"pepformRel")
#normalisation for protein abundance step
assay(pe[["pepformRel"]]) <- assay(pe[["pepformRel"]]) - assay(pe[["proteinRobust"]])[rowData(pe[["pepformRel"]])$Proteins,colnames(assay(pe[["pepformRel"]]))]
boxplot(assay(pe[["pepformRel"]]))
```

```{r}
filtering <- (rowSums(assay(pe[["pepformRel"]]) != 0, na.rm=TRUE)) > 0 & 
  (rowSds(assay(pe[["pepformRel"]]), na.rm=TRUE) > 1e-4) & 
          (rowSums(is.na(assay(pe[["pepformRel"]]))) < (ncol(assay(pe[["pepformRel"]]))-1))
pe[["pepformRel"]] <- pe[["pepformRel"]][filtering,]
pe <- zeroIsNA(pe, "pepformRel")
```



# DPTM

## ptm summarisatie

### Determine location of ptm

```{r}
fasta <- "human.fasta"
parsed_fasta <- read.fasta(file = fasta, seqtype = "AA", as.string = T)
```

  
```{r}
#modified sequence column contains _ that does not matter, but hinders location determining
rowData(pe[["pepformRel"]])$modified_sequence <- gsub("_", "", rowData(pe[["pepformRel"]])$Modified.sequence)
```

  

```{r}
get_ptm_location <- function(feature, data, fasta, mod_column = "Modifications", 
                             peptide_seq_column = "Sequence", mod_seq_column = "modified_sequence", 
                             protein_column = "Leading.razor.protein", split = ",", collapse = ", "){
  prot <- data[feature,][[protein_column]]
  pep_seq <- data[feature,][[peptide_seq_column]]
  mod_seq <- data[feature,][[mod_seq_column]]
  prot_seq <- fasta[[prot]][1]

  #find location of peptide in protein
  #add -1 here, so that the addition of the location later on is correct
  pep_location <- unlist(lapply(gregexpr(pattern = pep_seq, prot_seq), min)) - 1
  final_mod <- c()
  j <- mod_seq
  #go over the modifications inside the modified sequence
  for(mod in regmatches(mod_seq, gregexpr("\\(.*?\\)\\)", mod_seq, perl=T))[[1]]){
    #find location of modification in peptide
    mod_location <- unlist(lapply(gregexpr(pattern = mod, j, fixed = T), min))
    #get location in protein (-1 because else it gives you the location after because of the presence of the modification in the string)
    location <- mod_location + pep_location -1
    #add location to modification
    mod_ <- paste(mod, location)
    #save modification
    final_mod <- c(final_mod, mod_)
    #now remove current modification from the sequence, so that we can continue to the next mod
    str_sub(j, mod_location, nchar(mod)+mod_location-1) <- ""
  }
  return(paste(final_mod, collapse = collapse))
}
```

```{r}
rowData(pe[["pepformRel"]])$mod <- sapply(rownames(rowData(pe[["pepformRel"]])), function(x){
    get_ptm_location(x, rowData(pe[["pepformRel"]]), parsed_fasta)
}, USE.NAMES = F)
```

### Get PTM variable
In our case, a ptm is a unique protein - modification + location combination.

```{r}
#Add ptm variable = protein + modification 
#Do this only for peptidoforms with modifications
rowData(pe[["pepformRel"]])$ptm <- ifelse(rowData(pe[["pepformRel"]])$mod != "",
                                          #if there is mod, add ptm
                                               mapply(function(x, y) {
vector2 <- strsplit(y, ", ")[[1]]
result <- lapply(vector2, function(i) paste(x, i, sep = "-"))
final_result <- paste(unlist(result), collapse = ", ")
}, rowData(pe[["pepformRel"]])$Leading.razor.protein, rowData(pe[["pepformRel"]])$mod), 
#else add empty string
                                               "")
```

First, we need a vector of all of these unique ptms.

```{r}
#Get all unique PTMs
ptms <- unique(unlist(strsplit(rowData(pe[["pepformRel"]])$ptm, split = ", ", fixed = TRUE)))
```

For every unique ptm, we take all its associated peptidoforms and aggregate all its intensity values per sample into one intensity value per sample for that ptm.

```{r}
#For each ptm do
ptm_x_assay <- sapply(seq(1:length(ptms)), function(i){ 
  x <- ptms[i]
  #Get current protein and mod from ptm
  prot <- str_split(x, "-", 2)[[1]][1]
  current_ptm <- str_split(x, "-", 2)[[1]][2]
  #filter on that protein and on that mod to obtain all peptidoforms that correspond to the ptm
  pe_sub <- pe[["pepformRel"]][grepl(prot, rowData(pe[["pepformRel"]])$Leading.razor.protein, fixed = T),]
  ptm_sub <- pe_sub[grepl(current_ptm, rowData(pe_sub)$mod, fixed = T),]
  #Get intensity values of those peptidoforms
  y <- assay(ptm_sub)
  #And summarise them to 1 row of intensity values: 1 value per sample for that ptm
  ptm_y <- try(MsCoreUtils::robustSummary(y), silent = T)
  if (is(ptm_y, "try-error")){
    ptm_y <- rep(NA, ncol(y))}
  ptm_y
})
#Then we get the intensity assay on ptm level
ptm_x_assay <- t(ptm_x_assay)
rownames(ptm_x_assay) <- ptms
```

### Add to QFeatures object

Filter out ptms with all zero intensities

```{r}
print(paste(nrow(ptm_x_assay), "ptms before filtering"))
filtering <- (rowSums(ptm_x_assay != 0, na.rm=TRUE)) > 0 & 
  (rowSds(ptm_x_assay, na.rm=TRUE) > 1e-4) & 
          (rowSums(is.na(ptm_x_assay)) < (ncol(ptm_x_assay)-1)) & 
  (rowSums(is.finite(ptm_x_assay)) > 1)
ptm_x_assay <- ptm_x_assay[filtering,]
print(paste(nrow(ptm_x_assay), "ptms after filtering"))
```

```{r}
all(rownames(colData(pe)) == colnames(ptm_x_assay))
rowdata <- data.frame(ptm = rownames(ptm_x_assay))
rowdata$protein <- sapply(str_split(rowdata$ptm, pattern=" "),function(x) x[1])
ptm_y_assay <- SummarizedExperiment(assays=as.matrix(ptm_x_assay), rowData=rowdata, colData=colData(pe))
```


```{r}
pe <- addAssay(pe, ptm_y_assay, "ptmRel")
pe <- zeroIsNA(pe, "ptmRel")
```

# Simulations

### Peptidoform level

```{r warning=FALSE}
set.seed(745)
pvals <- matrix(NA,nrow=nrow(assay(pe[["pepformRel"]])),ncol=5)
adj_pvals <- matrix(NA,nrow=nrow(assay(pe[["pepformRel"]])),ncol=5)
fprs <- c()
L <- makeContrast(c("conditionA = 0"), parameterNames = c("(Intercept)", "conditionA"))

for (i in 1:5)
{
# if(i%%50==0){
#   print(i)
# }
#There are 26 samples
nSamp <- colData(pe) %>% rownames %>% unique() %>% length()
#We randomly choose 13 out of 26 samples
mocks <- sample(colData(pe) %>% rownames %>% unique(),nSamp%/%2)
colData(pe)$condition <- "B"
#those 9 get "A" assigned, rest gets B
colData(pe)[mocks,]$condition <- "A"
colData(pe)$condition <- relevel(as.factor(colData(pe)$condition), ref = "B")

colData(pe[["pepformRel"]]) <- colData(pe)

pe <- msqrob2::msqrob(object = pe, i = "pepformRel", formula = ~condition, robust=FALSE, overwrite = T)
pe <- hypothesisTest(object = pe, i = "pepformRel", contrast = L, overwrite = T)
pvals[,i] <- rowData(pe)[["pepformRel"]]$conditionA$pval
adj_pvals[,i] <- rowData(pe)[["pepformRel"]]$conditionA$adjPval
hist(pvals, breaks = 50, main = "Histogram of pvals - peptidoform level - no robust regression", cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5)
}

```



```{r warning=FALSE}
set.seed(745)
pvals <- matrix(NA,nrow=nrow(assay(pe[["pepformRel"]])),ncol=5)
adj_pvals <- matrix(NA,nrow=nrow(assay(pe[["pepformRel"]])),ncol=5)
L <- makeContrast(c("conditionA = 0"), parameterNames = c("(Intercept)", "conditionA"))

for (i in 1:5)
{
# if(i%%50==0){
#   print(i)
# }
#There are 18 samples
nSamp <- colData(pe) %>% rownames %>% unique() %>% length()
#We randomly choose 9 out of 18 samples
mocks <- sample(colData(pe) %>% rownames %>% unique(),nSamp%/%2)
colData(pe)$condition <- "B"
#those 9 get "A" assigned, rest gets B
colData(pe)[mocks,]$condition <- "A"
colData(pe)$condition <- relevel(as.factor(colData(pe)$condition), ref = "B")

colData(pe[["pepformRel"]]) <- colData(pe)

pe <- msqrob2::msqrob(object = pe, i = "pepformRel", formula = ~condition, robust=TRUE, overwrite = T)
pe <- hypothesisTest(object = pe, i = "pepformRel", contrast = L, overwrite = T)
pvals[,i] <- rowData(pe)[["pepformRel"]]$conditionA$pval
adj_pvals[,i] <- rowData(pe)[["pepformRel"]]$conditionA$adjPval
hist(pvals, breaks = 50, main = "Histogram of pvals - peptidoform level - robust regression")
}
```

### PTM level

```{r warning=FALSE}
set.seed(45)
pvals <- matrix(NA,nrow=nrow(assay(pe[["ptmRel"]])),ncol=5)
adj_pvals <- matrix(NA,nrow=nrow(assay(pe[["ptmRel"]])),ncol=5)
L <- makeContrast(c("conditionA = 0"), parameterNames = c("(Intercept)", "conditionA"))

for (i in 1:5)
{
# if(i%%50==0){
#   print(i)
# }
#There are 18 samples
nSamp <- colData(pe) %>% rownames %>% unique() %>% length()
#We randomly choose 9 out of 18 samples
mocks <- sample(colData(pe) %>% rownames %>% unique(),nSamp%/%2)
colData(pe)$condition <- "B"
#those 9 get "A" assigned, rest gets B
colData(pe)[mocks,]$condition <- "A"
colData(pe)$condition <- relevel(as.factor(colData(pe)$condition), ref = "B")

colData(pe[["ptmRel"]]) <- colData(pe)

pe <- msqrob2::msqrob(object = pe, i = "ptmRel", formula = ~condition, robust=FALSE, overwrite = T)
pe <- hypothesisTest(object = pe, i = "ptmRel", contrast = L, overwrite = T)
pvals[,i] <- rowData(pe)[["ptmRel"]]$conditionA$pval
adj_pvals[,i] <- rowData(pe)[["ptmRel"]]$conditionA$adjPval
hist(pvals, breaks = 50, main = "Histogram of pvals - no robust regression")
}
```



```{r warning=FALSE}
set.seed(45)
pvals <- matrix(NA,nrow=nrow(assay(pe[["ptmRel"]])),ncol=5)
adj_pvals <- matrix(NA,nrow=nrow(assay(pe[["ptmRel"]])),ncol=5)
L <- makeContrast(c("conditionA = 0"), parameterNames = c("(Intercept)", "conditionA"))

for (i in 1:5)
{
# if(i%%50==0){
#   print(i)
# }
#There are 18 samples
nSamp <- colData(pe) %>% rownames %>% unique() %>% length()
#We randomly choose 9 out of 18 samples
mocks <- sample(colData(pe) %>% rownames %>% unique(),nSamp%/%2)
colData(pe)$condition <- "B"
#those 9 get "A" assigned, rest gets B
colData(pe)[mocks,]$condition <- "A"
colData(pe)$condition <- relevel(as.factor(colData(pe)$condition), ref = "B")
colData(pe[["ptmRel"]]) <- colData(pe)

pe <- msqrob2::msqrob(object = pe, i = "ptmRel", formula = ~condition, robust=TRUE, overwrite = T)
pe <- hypothesisTest(object = pe, i = "ptmRel", contrast = L, overwrite = T)
pvals[,i] <- rowData(pe)[["ptmRel"]]$conditionA$pval
adj_pvals[,i] <- rowData(pe)[["ptmRel"]]$conditionA$adjPval
hist(pvals, breaks = 50, main = "Histogram of pvals - robust regression")
}

```

