---
title: "spike-in data msqrobPTM"
output:
  html_document:
    code_download: yes
    theme: cosmo
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
    highlight: tango
    number_sections: yes
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: '4'
---



```{r}
library(tidyverse)
library(MSstatsPTM)
library(QFeatures)
library(msqrob2)
library(gt)
library(seqinr)
library(data.table)
```

```{r, setup, include=F}
knitr::opts_knit$set(root.dir="C:/Users/Nina/OneDrive - UGent/Documenten/Doctoraat/msqrobPTM paper/spike-in dataset/")
```


```{r}
load("MSstatsPTM_Summarized.rda")
```


``` {r}
#In here PTM level and Protein level data
head(summary_df)
```


-> Get to wide format for QFeatures object

```{r}
#PTM dataset -> only the human proteins were used as global proteome (said in paper)
# -> check what is in here
human_fasta <- "homo_sapiens_all_20160725.fasta"
ecoli_fasta <- "Swissprot-Ecoli-K12.fasta"
human <- names(read.fasta(file = human_fasta, seqtype = "AA", as.string = T))
human <- sapply(strsplit(human, "|", fixed = T), function(x) x[[2]])
ecoli <- names(read.fasta(file = ecoli_fasta, seqtype = "AA", as.string = T))
ecoli <- sapply(strsplit(ecoli, "|", fixed = T), function(x) x[[2]])


#Get to wide format for msqrob
pepform_df <- pivot_wider(summary_df$PTM$ProteinLevelData, id_cols = c("Protein"),
                         names_from = "originalRUN", values_from = "LogIntensities")
pepform_df <- pepform_df %>% dplyr::rename(PROTEIN_PTM = Protein)
pepform_df$PROTEIN =  paste(sapply(pepform_df$PROTEIN_PTM, function (x) {str_split(x, "_")[[1]][1]}))
human_proteins <- pepform_df$PROTEIN %in% human
ecoli_proteins <- pepform_df$PROTEIN %in% ecoli
```

both in ptm level data as in protein level data there were already only human proteins

```{r}
head(pepform_df, 10)
```


Transform into QFeatures object

```{r}
#order dataframe by protein, for the normalisation step
pepform_df = pepform_df[order(pepform_df$PROTEIN),]
ecols <- grep("mix",names(pepform_df))
pepform_df <- pepform_df[,c(names(pepform_df)[-ecols],sort(names(pepform_df)[ecols]))]
ecols <- grep("mix", colnames(pepform_df))
pe <- readQFeatures(pepform_df,ecol= ecols,name="ptmRaw", fnames = "PROTEIN_PTM")
```


```{r}
colData(pe)$mix <- sapply(str_split(rownames(colData(pe)), "-"), function(x) x[1])
colData(pe)$mix <- as.factor(colData(pe)$mix)
colData(pe)$biorep <- rownames(colData(pe))
colData(pe)
```

```{r}
#Get to wide format for msqrob
protein_df <- pivot_wider(summary_df$PROTEIN$ProteinLevelData, id_cols = c("Protein"),
                         names_from = "originalRUN", values_from = "LogIntensities")
ecols <- grep("mix", colnames(protein_df))
protein_df <- protein_df[,c(names(protein_df)[-ecols],sort(names(protein_df)[ecols]))]
ecols <- grep("mix", colnames(protein_df))
prot <- readQFeatures(protein_df,ecol= ecols,name="protein",fnames="Protein")
colData(prot) <- colData(pe)
```




# Preprocessing

```{r}
##PTM
print(nrow(pe[["ptmRaw"]]))
rowData(pe[["ptmRaw"]])$nNonZero <- rowSums(assay(pe[["ptmRaw"]]) > 0, na.rm = T)
pe <- zeroIsNA(pe, i = "ptmRaw")
pe <- pe[rowData(pe[["ptmRaw"]])$nNonZero>2,,]
print(nrow(pe[["ptmRaw"]]))
```

```{r}
##Protein 
prot <- zeroIsNA(prot, i = "protein")
rowData(prot[["protein"]])$nNonZero <- rowSums(assay(prot[["protein"]]) > 0, na.rm = T)
nrow(prot[["protein"]])
#prot <- logTransform(prot, base = 2, i = "proteinRaw", name = "protein")
prot <- prot[rowData(prot[["protein"]])$nNonZero>2,,]
nrow(prot[["protein"]])
colData(prot[["protein"]]) <- colData(prot)
```



# Normalize PTM log intensity by protein log intensity

```{r}
peL <- pe[,,1]
#Only take ptms that have a parent protein
ptmWithProteinEvalue <- rowData(peL[["ptmRaw"]])$PROTEIN %in% rownames(prot[["protein"]])
peL <- peL[ptmWithProteinEvalue,,]
peL <- addAssay(peL,peL[["ptmRaw"]],"ptmRel")
#normalisation for protein abundance step
assay(peL[["ptmRel"]]) <- assay(peL[["ptmRel"]]) - assay(prot[["protein"]])[rowData(peL[["ptmRel"]])$PROTEIN,colnames(assay(peL[["ptmRel"]]))]
boxplot(assay(peL[["ptmRel"]]))
```

# Median normalization 

```{r}
peL <- normalize(peL, i = "ptmRel", method = "center.median", name = "ptmRelNorm")
boxplot(assay(peL[["ptmRelNorm"]]))
```

```{r}
peL
```

ptmRaw = raw ptm intensities (no normalisations), but only for ptms that have a parent protein
ptmRel = relative abundance, normalised for protein
ptmRelNorm = ptmRel + median centring


# Analysis

```{r}
peL <- msqrob(peL,i="ptmRel",formula=~mix,robust=FALSE)
#ptmRelNorm is nog met center median normalisatie
peL <- msqrob(peL,i="ptmRelNorm",formula=~mix,robust=FALSE)

L <- makeContrast(c("mixmix2 = 0",
                    "mixmix3 = 0",
                    "mixmix4 = 0",
                    "mixmix4 - mixmix2 = 0",
                    "mixmix3 - mixmix2 = 0",
                    "mixmix4 - mixmix3 = 0"), 
                  parameterNames = rowData(peL[["ptmRelNorm"]])$msqrobModels[[2]] %>%
                  getCoef %>%
                  names)

peL <- hypothesisTest(object = peL, i = "ptmRelNorm", contrast = L, overwrite = T)
peL <- hypothesisTest(object = peL, i = "ptmRel", contrast = L, overwrite = T)
rowData(peL[["ptmRelNorm"]])$heavy <- grepl("heavy", rownames(peL[["ptmRelNorm"]]))
rowData(peL[["ptmRel"]])$heavy <- grepl("heavy", rownames(peL[["ptmRel"]]))
```

# Evaluation

```{r}
tprFdp <- function(pval, tp, adjPval) {
    ord <- order(pval)
    return(data.frame(
        pval = pval[ord],
        adjPval = adjPval[ord],
        tpr = cumsum(tp[ord]) / sum(tp),
        fpr = cumsum(!tp[ord]) / sum(!tp),
        fdp = cumsum(!tp[ord]) / 1:length(tp)
    ))
}

tprFdpPtmRel <- lapply(colnames(L),function(contrast) 
  tprFdp(
    rowData(peL[["ptmRel"]])[[contrast]]$pval,
    rowData(peL[["ptmRel"]])$heavy,
    rowData(peL[["ptmRel"]])[[contrast]]$adjPval
    )
  )

tprFdpPtmRelNorm <- lapply(colnames(L),function(contrast) 
  tprFdp(
    rowData(peL[["ptmRelNorm"]])[[contrast]]$pval,
    rowData(peL[["ptmRelNorm"]])$heavy,
    rowData(peL[["ptmRelNorm"]])[[contrast]]$adjPval
    )
  )
```

# MSstats Emulation 

```{r}
peL <- normalize(peL, i = "ptmRaw", method = "center.median", name = "ptmNorm")
prot <- normalize(prot, i = "protein", method = "center.median", name = "proteinNorm")
```

```{r}
peL
```

ptmRaw = raw ptm intensities (no normalisations), but only for ptms that have a parent protein
ptmNorm = median centered ptm intensities, not corrected for protein abundance
ptmRel = relative abundance, normalised for protein
ptmRelNorm = ptmRel + median centering

```{r, fig.width=9}
boxplot(assay(peL[["ptmRaw"]]))
boxplot(assay(peL[["ptmNorm"]]))
boxplot(assay(prot[["protein"]]))
boxplot(assay(prot[["proteinNorm"]]))
```

```{r}
peL <- msqrob(peL,i="ptmNorm",formula=~mix,robust=FALSE,overwrite=TRUE)
peL <- hypothesisTest(object = peL, i = "ptmNorm", contrast = L, overwrite = T)
peL <- msqrob(peL,i="ptmRaw",formula=~mix,robust=FALSE,overwrite=TRUE)
peL <- hypothesisTest(object = peL, i = "ptmRaw", contrast = L, overwrite = T)

prot <- msqrob(prot,i="proteinNorm",formula=~mix,robust=FALSE,overwrite=TRUE)
prot <- hypothesisTest(object = prot, i = "proteinNorm", contrast = L, overwrite = T)
prot <- msqrob(prot,i="protein",formula=~mix,robust=FALSE,overwrite=TRUE)
prot <- hypothesisTest(object = prot, i = "protein", contrast = L, overwrite = T)
```

Msstats emulation with EB

```{r}
msstatsResEB <- lapply(colnames(L),function(contrast) 
{
  resPtm <- rowData(peL[["ptmRaw"]])[[contrast]]
  resProt <- (rowData(prot[["protein"]])[[contrast]])[rowData(peL[["ptmRaw"]])$PROTEIN,]
  resPtm$logFC <- resPtm$logFC-resProt$logFC
  s2 <- resPtm$se^2
  s2prot <- resProt$se^2
  resPtm$se <- sqrt(s2+s2prot)
  numer <- (s2+s2prot)^2
  denom <- (s2^2/resPtm$df+s2prot^2/resProt$df)
  resPtm$df <- numer/denom
  resPtm$t <- resPtm$logFC/resPtm$se
  resPtm$pval <- 2*pt(abs(resPtm$t),resPtm$df,lower.tail=FALSE)
  resPtm$adjPval <- p.adjust(resPtm$pval,method="fdr")
  return(resPtm)
})
names(msstatsResEB) <- colnames(L)
```

msqrob does EB by default while MsStatsPTM does not use empirical Bayes. 
Code to recalculate se en df 

```{r}
msstatsRes <- lapply(colnames(L),function(contrast) 
{
  resPtm <- rowData(peL[["ptmRaw"]])[[contrast]]
  resPtm$df <- sapply(rowData(peL[["ptmRaw"]])$msqrobModels,getDF)
  resPtm$se <- resPtm$se *sapply(rowData(peL[["ptmRaw"]])$msqrobModels,getSigma)/sapply(rowData(peL[["ptmRaw"]])$msqrobModels,getSigmaPosterior)
  resProt <- rowData(prot[["protein"]])[[contrast]]
  resProt$se <- resProt$se *sapply(rowData(prot[["protein"]])$msqrobModels,getSigma)/sapply(rowData(prot[["protein"]])$msqrobModels,getSigmaPosterior)
  resProt$df <- sapply(rowData(prot[["protein"]])$msqrobModels,getDF)
  resProt <- resProt[rowData(peL[["ptmRaw"]])$PROTEIN,]
  #resProt <- (rowData(prot[["protein"]])[[contrast]])[rowData(peL[["ptmRaw"]])$PROTEIN,]
  resPtm$logFC <- resPtm$logFC-resProt$logFC
  s2 <- resPtm$se^2
  s2prot <- resProt$se^2
  resPtm$se <- sqrt(s2+s2prot)
  numer <- (s2+s2prot)^2
  denom <- (s2^2/resPtm$df+s2prot^2/resProt$df)
  resPtm$df <- numer/denom
  resPtm$t <- resPtm$logFC/resPtm$se
  resPtm$pval <- 2*pt(abs(resPtm$t),resPtm$df,lower.tail=FALSE)
  resPtm$adjPval <- p.adjust(resPtm$pval,method="fdr")
  return(resPtm)
})
names(msstatsRes) <- colnames(L)
```



```{r}
load(file="MSstatsPTM_Summarized.rda")#raw_input_data
model_ptm <- groupComparisonPTM(summary_df, data.type = "LabelFree")

resMsStatsNative <- model_ptm$ADJUSTED.Model
resMsStatsNative$Label <- as.factor(resMsStatsNative$Label)
levels(resMsStatsNative$Label) <- c("mixmix2","mixmix3 - mixmix2","mixmix4 - mixmix2","mixmix3","mixmix4 - mixmix3","mixmix4")
resMsStatsNative <- lapply(colnames(L), function(contrast) resMsStatsNative %>% filter(Label==contrast))
names(resMsStatsNative) <- colnames(L)
```

```{r}
rowData(peL[["ptmRaw"]])$heavy <- grepl("heavy", rownames(peL[["ptmRaw"]]))
rowData(peL[["ptmNorm"]])$heavy <- grepl("heavy", rownames(peL[["ptmNorm"]]))

tprFdpPtmRaw <- lapply(colnames(L),function(contrast) 
  tprFdp(
    rowData(peL[["ptmRaw"]])[[contrast]]$pval,
    rowData(peL[["ptmRaw"]])$heavy,
    rowData(peL[["ptmRaw"]])[[contrast]]$adjPval
    )
  )


tprFdpPtmNorm <- lapply(colnames(L),function(contrast) 
  tprFdp(
    rowData(peL[["ptmNorm"]])[[contrast]]$pval,
    rowData(peL[["ptmNorm"]])$heavy,
    rowData(peL[["ptmNorm"]])[[contrast]]$adjPval
    )
)

tprFdpPtmNormNoEB <- lapply(colnames(L),function(contrast) 
{
  datah <- rowData(peL[["ptmNorm"]])[[contrast]]
  datah$se<- datah$se*sapply(rowData(peL[["ptmNorm"]])$msqrobModels,getSigma)/sapply(rowData(peL[["ptmNorm"]])$msqrobModels,getSigma)
  datah$df<-sapply(rowData(peL[["ptmNorm"]])$msqrobModels,getDF)
  datah$t <- datah$logFC/datah$se
  datah$pval <- pt(abs(datah$t),datah$df,lower.tail = FALSE)*2
  datah$adjPval<-p.adjust(datah$pval,method="fdr")
  return(tprFdp(
    datah$pval,
    rowData(peL[["ptmNorm"]])$heavy,
    datah$adjPval
    )
    )
})

tprFdpPtmRawMsStats <- lapply(colnames(L),function(contrast) 
  {
  tprFdp(
    msstatsRes[[contrast]]$pval,
    grepl("heavy", rownames(msstatsRes[[contrast]])),
    msstatsRes[[contrast]]$adjPval
    )
  })

tprFdpPtmRawMsStatsEB <- lapply(colnames(L),function(contrast) 
  {
  tprFdp(
    msstatsResEB[[contrast]]$pval,
    grepl("heavy", rownames(msstatsResEB[[contrast]])),
    msstatsResEB[[contrast]]$adjPval
    )
  })

tprFdpMsstatsNative <- lapply(colnames(L),function(contrast) 
  {
  tprFdp(
    resMsStatsNative[[contrast]]$pvalue,
    grepl("heavy", resMsStatsNative[[contrast]]$Protein),
    resMsStatsNative[[contrast]]$adj.pvalue
    )
  })

```

### ROC curves

```{r}
names(tprFdpPtmRaw) <- colnames(L)
names(tprFdpPtmNorm) <- colnames(L)
names(tprFdpPtmNormNoEB) <- colnames(L)
names(tprFdpPtmRel) <- colnames(L)
names(tprFdpPtmRelNorm) <- colnames(L)
names(tprFdpPtmRawMsStats) <- colnames(L)
names(tprFdpPtmRawMsStatsEB) <- colnames(L)
names(tprFdpMsstatsNative) <- colnames(L)

rocPlots <- lapply(colnames(L),
function(contrast)
{
  hlp <- rbind(
    cbind(tprFdpPtmRaw[[contrast]], method = "DPA"),
    cbind(tprFdpPtmNorm[[contrast]], method = "DPA-norm"),
    cbind(tprFdpPtmRel[[contrast]], method = "msqrob"),
    cbind(tprFdpPtmRelNorm[[contrast]], method = "msqrob-norm"),
    cbind(tprFdpPtmRawMsStats[[contrast]], method = "msstatsEmul"),
    cbind(tprFdpPtmRawMsStatsEB[[contrast]], method = "msstatsEmulEB"),
    cbind(tprFdpMsstatsNative[[contrast]], method = "msstatsNative")
    )
hlp %>%
    ggplot(aes(x = fpr, y = tpr, color = method)) +
    geom_path() +
    ggtitle(contrast)
})
rocPlots
```

```{r, fig.height = 11, fig.width=15}
names(tprFdpPtmRaw) <- colnames(L)
names(tprFdpPtmNorm) <- colnames(L)
names(tprFdpPtmNormNoEB) <- colnames(L)
names(tprFdpPtmRel) <- colnames(L)
names(tprFdpPtmRelNorm) <- colnames(L)
names(tprFdpPtmRawMsStats) <- colnames(L)
names(tprFdpPtmRawMsStatsEB) <- colnames(L)
names(tprFdpMsstatsNative) <- colnames(L)

hlp <- sapply(colnames(L),
function(contrast)
{
  tmp <- rbind(
    cbind(tprFdpPtmRaw[[contrast]], method = "DPA-NonNorm", contrast = contrast),
    cbind(tprFdpPtmNorm[[contrast]], method = "DPA", contrast = contrast),
    #cbind(tprFdpPtmRel[[contrast]], method = "msqrob2PTM-NonNorm", contrast = contrast),
    cbind(tprFdpPtmRelNorm[[contrast]], method = "msqrob2PTM", contrast = contrast),
    #cbind(tprFdpPtmRawMsStats[[contrast]], method = "msstatsEmul", contrast = contrast),
    #cbind(tprFdpPtmRawMsStatsEB[[contrast]], method = "msstatsEmulEB", contrast = contrast),
    cbind(tprFdpMsstatsNative[[contrast]], method = "MSstatsPTM", contrast = contrast)
    )
  tmp
}, simplify = F)
hlp <- do.call("rbind", hlp)
hlp %>%
    ggplot(aes(x = fpr, y = tpr, color = method)) +
    geom_path(size= 1) +
    scale_colour_manual(values = c("#7CAE00", "#D62728", "#C77CFF",
                                   "#FAA43A"
                                   )) +
    facet_wrap(~contrast) +
    theme(legend.position = "bottom",
          axis.text=element_text(size=11),
          axis.title=element_text(size=15),
          strip.text.x = element_text(size = 18),
          strip.text.y = element_text(size = 18),
          legend.title = element_text(size=14), 
          legend.text = element_text(size=14)) 
```

tpr-fdp curves

```{r, fig.height = 9, fig.width=13}
hlp %>%
    ggplot(aes(x = fdp, y = tpr, color = method)) +
    geom_path(size= 1) +
    scale_colour_manual(values = c("#D62728", "#7CAE00", 
                                   "#C77CFF", "#FAA43A", "#4D4D4D",
                                   "#00BFC4", "#F0E442")) +
    facet_wrap(~contrast) +
    theme(legend.position = "bottom",
          axis.text=element_text(size=11),
          axis.title=element_text(size=15),
          strip.text.x = element_text(size = 18),
          strip.text.y = element_text(size = 18),
          legend.title = element_text(size=14), 
          legend.text = element_text(size=14))
```


# Boxplots of fold changes


```{r, echo = FALSE, fig.width = 8}
summary_together_df <- rbind(rowData(peL[["ptmRelNorm"]])$mixmix2 %>% rownames_to_column("Protein") %>% mutate(Contrast = "mix2 vs mix1"), 
                               rowData(peL[["ptmRelNorm"]])$mixmix3 %>% rownames_to_column("Protein") %>% mutate(Contrast = "mix3 vs mix1"),
                               rowData(peL[["ptmRelNorm"]])$mixmix4 %>% rownames_to_column("Protein") %>% mutate(Contrast = "mix4 vs mix1"),
                               rowData(peL[["ptmRelNorm"]])[["mixmix4 - mixmix2"]] %>% rownames_to_column("Protein") %>% mutate(Contrast = "mix4 vs mix2"),
                               rowData(peL[["ptmRelNorm"]])[["mixmix3 - mixmix2"]] %>% rownames_to_column("Protein") %>% mutate(Contrast = "mix3 vs mix2"),
                               rowData(peL[["ptmRelNorm"]])[["mixmix4 - mixmix3"]] %>% rownames_to_column("Protein") %>% mutate(Contrast = "mix4 vs mix3"))
summary_together_df <- summary_together_df %>% mutate(labeling = ifelse(grepl("heavy", summary_together_df$Protein), "heavy", "light"))
```

#### Log Fold Change - Spike-in Peptides

Lastly, the plot below looks at the distribution spike-in peptides log fold changes. The expected log Fold Change is indicated by the red `X` marks.

```{r, echo = FALSE, fig.width = 10, fig.height = 5}
summary_together_df <- summary_together_df %>% 
                        mutate(trueLog2FC = case_when(
                          Contrast == "mix2 vs mix1" ~ -1,
                          Contrast == "mix3 vs mix1" ~ 1,
                          Contrast == "mix4 vs mix1" ~ 0,
                          Contrast == "mix3 vs mix2" ~ 2,
                          Contrast == "mix4 vs mix2" ~ 1,
                          Contrast == "mix4 vs mix3" ~ -1
                        ))

summary_together_df %>% filter(labeling == "heavy") %>% ggplot() + geom_boxplot(aes(x = Contrast, y = logFC), fill = "lightsteelblue4") + 
  geom_point(aes(y = trueLog2FC, x = Contrast), color = "red", shape = 4, size = 3) + 
  labs(title = "MsqrobPTM Fold Change of Spike-in Peptides", x = "Comparison") + theme(axis.text.x = element_text(angle = 30, vjust = .5))
 
```

```{r}
summary_together_df %>% mutate(LogFClight = 0) %>% filter(labeling == "light") %>% 
  ggplot() + geom_boxplot(aes(x = Contrast, y = logFC), fill = "lightsteelblue4") + 
  geom_point(aes(y = LogFClight, x = Contrast), color = "red", shape = 4, size = 3) + 
  labs(title = "MsqrobPTM Fold Change of background Peptides", x = "Comparison") + theme(axis.text.x = element_text(angle = 30, vjust = .5))
```


```{r, echo = FALSE, fig.width = 10, fig.height = 7}
load("MSstatsPTM_Model.rda")

msstats <- model_df$ADJUSTED.Model %>% 
  select(-c("GlobalProtein", "protadj")) %>% mutate(model = "MSstatsPTM") 
msstats$group = paste(msstats$Label, msstats$model)
colnames(msstats) <- c("Protein", "Contrast", "logFC", "se", "t", "df", 
                             "pval", "adjPval", "labeling", "trueLog2FC", 
                             "contrast_truth", "model", "group")

summary_together_df$model = "MsqrobPTM"
summary_together_df$group = paste(summary_together_df$Contrast, summary_together_df$model)

summary_together_df <- summary_together_df %>% mutate(
  contrast_truth = as.factor(case_when(Contrast == 'mix2 vs mix1' ~ "LogFC: -1 (mix2-mix1)",
                             Contrast == 'mix3 vs mix1' ~ "LogFC: 1 (mix3-mix1)",
                             Contrast == 'mix4 vs mix1' ~ "LogFC: 0 (mix4-mix1)",
                             Contrast == 'mix3 vs mix2' ~ "LogFC: 2 (mix3-mix2)",
                             Contrast == 'mix4 vs mix2' ~ "LogFC: 1 (mix4-mix2)",
                             Contrast == 'mix4 vs mix3' ~ "LogFC: -1 (mix4-mix3)")
))

summary_together_df$labeling <- as.factor(summary_together_df$labeling)

rbind(msstats, summary_together_df) %>% filter(labeling == "heavy") %>% ggplot() + geom_boxplot(aes(x = group, y = logFC, fill = model)) + 
  scale_fill_manual(values = c("MsqrobPTM" = "deeppink2", "MSstatsPTM" = "dodgerblue2")) +
  geom_point(aes(y = trueLog2FC, x = group), color = "white", shape = 4, size = 4) + 
  labs(title = "MsqrobPTM and MSstatsPTM Fold Change of Spike-in Peptides", x = "Comparison")  + 
  theme(axis.text.x = element_text(angle = 90, size = 11.5),
        axis.text.y = element_text(size = 10),
        legend.text=element_text(size=12),
        axis.title.y = element_text(size = 14),
        axis.title.x = element_text(size = 14)) 
 
```

```{r, echo = FALSE, fig.width = 10, fig.height = 7}
load("MSstatsPTM_Model.rda")


summary_together_df$model = "MsqrobPTM"
summary_together_df$group = paste(summary_together_df$Contrast, summary_together_df$model)

rbind(msstats, summary_together_df) %>% filter(labeling == "light") %>% ggplot() + geom_boxplot(aes(x = group, y = logFC, fill = model)) + 
  scale_fill_manual(values = c("MsqrobPTM" = "deeppink2", "MSstatsPTM" = "dodgerblue2")) +
  geom_point(aes(y = 0, x = group), color = "navajowhite", shape = 4, size = 4) + 
  labs(title = "MsqrobPTM and MSstatsPTM Fold Change of background Peptides", x = "Comparison")  + 
  theme(axis.text.x = element_text(angle = 90, size = 11.5),
        axis.text.y = element_text(size = 10),
        legend.text=element_text(size=12),
        axis.title.y = element_text(size = 14),
        axis.title.x = element_text(size = 14)) 
 
```

#### Calculating background peptide percentages

```{r, echo=FALSE}

light_pep <- summary_together_df %>% filter(labeling == "light")

light_sign <- summary_together_df %>% filter(labeling == "light" & adjPval < .05)

nrow(light_sign) / nrow(light_pep)

```

# Volcano plots

```{r, fig.width=15, fig.height=7}
# Volcano plot
summary_together_df %>% arrange(desc(labeling)) %>% 
  filter(abs(logFC) < Inf) %>% # & Label %in% c("mix1-mix3", "mix1-mix4", "mix2-mix3", "mix2-mix4")
  ggplot(aes(logFC, -log10(adjPval))) + 
  geom_point(aes(color = labeling, alpha = labeling), size = 2.5) + 
  scale_color_manual(values = c("light" = "grey77", "heavy" = "hotpink")) + 
  scale_alpha_manual(values = c("light" = 0.2, "heavy" = 1)) + 
  geom_vline(xintercept = 0) + 
  geom_vline(aes(xintercept = trueLog2FC),color = "springgreen4", linetype= "dashed") +
  facet_grid(~contrast_truth) + 
  theme_bw() + 
  theme(axis.text.x = element_text(size = 16), 
       axis.text.y = element_text(size = 16), 
       legend.text=element_text(size=16),
       axis.title.y = element_text(size = 18, vjust=2),
       axis.title.x = element_text(size = 18),
       title = element_text(size = 20),
       strip.text = element_text(size = 15),
       legend.title =  element_blank(),
       legend.position = "bottom",
       legend.box.margin=margin(-10,-10,-10,-10)) + 
  scale_y_continuous(limits = c(0,3.5)) +
  scale_x_continuous(limits = c(-10,10)) +
  guides(colour = guide_legend(override.aes = list(size=5))) +
  labs(title = "Proposed Method VolcanoPlot")
```

```{r}
msstats = msstats %>% mutate(
  contrast_truth = case_when(
    contrast_truth == "mix2 vs mix1 (-1)" ~ "LogFC: -1 (mix2-mix1)",
    contrast_truth == "mix3 vs mix1 (1)" ~ "LogFC: 1 (mix3-mix1)",
    contrast_truth == "mix4 vs mix1 (0)" ~ "LogFC: 0 (mix4-mix1)",
    contrast_truth == "mix3 vs mix2 (2)" ~ "LogFC: 2 (mix3-mix2)",
    contrast_truth == "mix4 vs mix2 (1)" ~ "LogFC: 1 (mix4-mix2)",
    contrast_truth == "mix4 vs mix3 (-1)" ~ "LogFC: -1 (mix4-mix3)"
  )
)
```


```{r, fig.width=15, fig.height=9}
# Volcano plot
rbind(msstats, summary_together_df)  %>% arrange((labeling)) %>% 
  filter(abs(logFC) < Inf) %>% # & Label %in% c("mix1-mix3", "mix1-mix4", "mix2-mix3", "mix2-mix4")
  ggplot(aes(logFC, -log10(adjPval))) + 
  geom_point(aes(color = labeling, alpha = labeling), size = 2.5) + 
  scale_color_manual(values = c("light" = "grey77", "heavy" = "hotpink")) + 
  scale_alpha_manual(values = c("light" = 0.2, "heavy" = 1)) + 
  geom_vline(xintercept = 0) + 
  geom_vline(aes(xintercept = trueLog2FC),color = "springgreen4", linetype= "dashed", size = 1.2) +
  facet_grid(model~contrast_truth) + 
  theme_bw() + 
  theme(axis.text.x = element_text(size = 16), 
       axis.text.y = element_text(size = 16), 
       legend.text=element_text(size=16),
       axis.title.y = element_text(size = 18, vjust=2),
       axis.title.x = element_text(size = 18),
       title = element_text(size = 20),
       strip.text = element_text(size = 15),
       legend.title =  element_blank(),
       legend.position = "bottom",
       legend.box.margin=margin(-10,-10,-10,-10)) + 
  scale_y_continuous(limits = c(0,3.5)) +
  scale_x_continuous(limits = c(-10,10)) +
  guides(colour = guide_legend(override.aes = list(size=5))) +
  labs(title = "VolcanoPlot light and heavy peptides")
```

# Lowlevel checks 

```{r}
rocPlotsMsStats <- lapply(colnames(L),
function(contrast)
{
  dataEmul <- na.exclude(msstatsRes[[contrast]])
  dataEmulEB <- na.exclude(msstatsResEB[[contrast]])
  dataNative <- resMsStatsNative[[contrast]]%>% filter(Protein %in% rownames(dataEmul))
  hlp <- rbind(
    cbind(
      tprFdp(
        dataEmul$pval,
        grepl("heavy", rownames(dataEmul)),
        dataEmul$adjPval
        ),
      method ="msstatEmul"
    ),
        cbind(
      tprFdp(
        dataEmulEB$pval,
        grepl("heavy", rownames(dataEmulEB)),
        dataEmulEB$adjPval
        ),
      method ="msstatEmulEB"
    ),
        cbind(
      tprFdp(
        dataNative$pvalue,
        grepl("heavy", dataNative %>% pull(Protein)),
        dataNative$adj.pvalue
        ),
      method ="msstatNative"
    )
    )

  hlp %>%
    ggplot(aes(x = fpr, y = tpr, color = method,size=method)) +
    geom_line() +
    scale_size_manual(values = c("msstatEmul" = 2, "msstatEmulEB" = 1, "msstatNative" = 1)) +
    ggtitle(contrast)
})

rocPlotsMsStats
```
